### 2024-11-26 paper/paper.qmd
Upload model part

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

## Estimand

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model {#sec-model}

In this paper, a Bayesian logistic model will be utilized to understand how factors such as application type, submission date, and planning district influence the decisions of the application committee. This model estimates the probability of an application being approved (P(Approval)) based on the given predictors.

The goal of this modeling strategy is twofold: (1) to estimate how each predictor affects the likelihood of approval, and (2) to quantify the uncertainty associated with these estimates. This approach is well-suited for incorporating prior knowledge while addressing the hierarchical nature of the predictors, such as geographic regions.

Here we briefly describe the Bayesian analysis model used to investigate, and include justification for model and the variables, as well as discuss underlying assumptions, potential limitations, software used to implement the model, and evidence of model validation and checking.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Combining all the components, the complete model can be expressed as:

\begin{equation}
\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{application type} + \beta_2 \times \text{submission date} + \beta_3 \times \text{planning district}
\end{equation}

- $\hat{p}$: Decision outcome (1 = approval, 0 = refusal).
- $\beta_0$: Intercept, representing the baseline log-odds of approval for the reference levels of the predictors.
- $\beta_1$: Coefficients for application_type (dummy-coded, with CO as the reference category).
- $\beta_2$: Coefficient for date (measured as the year of submission to capture temporal trends).
- $\beta_3$: Coefficients for planning_district (dummy-coded, with North York as the reference category).

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a ... relationship between ... In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

### Model Implementation

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 scripts/06-model_data.R

@@ -11,19 +11,29 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(lintr)
+library(styler)
+library(here)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
+
+# Ensure categorical variables are treated as factors
+analysis_data <- analysis_data %>%
+  mutate(
+    application_type = as.factor(application_type),
+    planning_district = as.factor(planning_district)
+  )
+
+#### Model data ####
+first_model <- 
   stan_glm(
-    formula = flying_time ~ length + width,
+    formula = decision ~ application_type + date + planning_district,
     data = analysis_data,
-    family = gaussian(),
+    family = binomial(link = "logit"), # Logistic regression
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
     seed = 853
   )
 
@@ -35,3 +45,6 @@ saveRDS(
 )
 
 
+#### Lint and style the code ####
+lint(filename = here("scripts/00-simulate_data.R"))
+style_file(path = here("scripts/00-simulate_data.R"))

### 2024-11-26 paper/paper.qmd
Upload model justification and assumption

@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+In the Bayesian logistic regression model, the response variable $y$ follows a Bernoulli distribution, reflecting its binary nature.
+Coefficients $\beta_k$ (where k=0,1,2,3) of every predictor variables follow a Normal distribution.
+Specifically:
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|p_i &\sim \mbox{Bern}(p_i) \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5)
 \end{align}
@@
 \begin{equation}
-\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{application type} + \beta_2 \times \text{submission date} + \beta_3 \times \text{planning district}
+\log\left(\frac{p_i}{1 - p_i}\right) = \beta_0 + \beta_1 \times \text{application type} + \beta_2 \times \text{submission date} + \beta_3 \times \text{planning district}
 \end{equation}
 
-- $\hat{p}$: Decision outcome (1 = approval, 0 = refusal).
+- $p_i$: the probability for the binary outcome variable to be 1 (1 = approval, 0 = refusal).
 - $\beta_0$: Intercept, representing the baseline log-odds of approval for the reference levels of the predictors.
-- $\beta_1$: Coefficients for application_type (dummy-coded, with CO as the reference category).
-- $\beta_2$: Coefficient for date (measured as the year of submission to capture temporal trends).
-- $\beta_3$: Coefficients for planning_district (dummy-coded, with North York as the reference category).
+- $\beta_1$: Coefficients for application type, dummy-coded with CO as the reference category.
+- $\beta_2$: Coefficient for date, measured as the year of application to capture temporal trends.
+- $\beta_3$: Coefficients for planning district, dummy-coded with North York as the reference category.
+
+In the above Bayesian logistic regression model, the coefficients $\beta = {\beta_0, \beta_1, \beta_2, \beta_3}$ are treated as random variables, each following a specified prior distribution.
+These priors represent our initial beliefs about the plausible values of the coefficients before observing the data.
+In this model, each $\beta_k$ (where k=0,1,2,3) is assigned a normal distribution, with a zero mean and 2.5 standard deviation.
+The choice of these parameters reflects a weakly informative prior, centered around zero to express no strong prior belief about the direction or magnitude of the effects, while allowing for reasonable variability in the coefficient estimates.
+
+The normal distribution is particularly suitable for these priors because it reflects a belief that most effects are likely to be small or moderate, centered around zero, while allowing for deviations in either direction.
+The standard deviation of 2.5 is a practical choice for weakly informative priors, offering a balance between constraining the parameters and allowing the data to influence the estimates.
+This ensures that the model avoids overfitting, particularly when data is sparse or multicollinearity exists among predictors.
+
+After observing the data, the posterior distributions of $\beta$ are obtained by combining these priors with the likelihood of the observed outcomes.
+The resulting posterior distributions provide updated beliefs about the coefficients, incorporating both prior knowledge and evidence from the data.
+This approach allows for robust inference, offering a clear quantification of parameter uncertainty and supporting more nuanced interpretations of the predictors' effects.
 
@@
 
-We expect a ... relationship between ... In particular...
+The Bayesian logistic regression model is an appropriate choice for analyzing how application type, submission date, and planning district influence the likelihood of application approval.
+This choice is driven by the binary nature of the response variable, the model's ability to incorporate prior knowledge, and its probabilistic framework, which quantifies uncertainty more effectively than frequentist alternatives.
+Logistic regression naturally constrains predicted probabilities to the [0,1] interval, ensuring interpretability as likelihoods of approval, while the Bayesian approach enhances flexibility and provides richer insights through posterior distributions.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Compared to traditional generalized linear models (GLM), the Bayesian approach allows for the inclusion of prior beliefs through weakly informative priors ($N(0, 2.5)$) for all coefficients.
+These priors help stabilize the model, particularly in cases where data is sparse or multicollinear, without overwhelming the contribution of the observed data.
+Additionally, Bayesian methods yield full posterior distributions rather than single-point estimates, allowing for a clearer understanding of parameter uncertainty and a better capacity to incorporate uncertainty into decision-making processes.
+These advantages make the Bayesian logistic regression model preferable to a standard GLM, which provides point estimates and relies on asymptotic approximations for inference.
+
+The model also stands out against simpler approaches, such as simple linear regression (SLR).
+While SLR might be used for binary outcomes in certain contexts, it is theoretically inappropriate because it does not constrain predictions to the [0,1] interval.
+This can result in nonsensical predicted probabilities outside this range.
+Moreover, SLR assumes a linear relationship between predictors and the response variable, which is unsuitable for binary outcomes.
+The logit transformation in logistic regression, by contrast, ensures a proper probabilistic framework while preserving interpretability in terms of log-odds.
+
+The predictors included in the model further justify the choice of Bayesian logistic regression.
+Application type captures fundamental differences in how various categories of applications, such as minor variances and consents, might affect approval outcomes.
+Submission date is treated as a continuous variable, reflecting temporal trends without arbitrary grouping, while planning district accounts for geographic variability through categorical indicators.
+These predictors are naturally suited to a logistic framework, and the Bayesian approach accommodates any inherent variability in their effects.
+
+In summary, Bayesian logistic regression provides a robust, clarified, and flexible framework for analyzing the factors influencing application decisions.
+Its advantages over GLM and SLR include the ability to incorporate prior information, quantify uncertainty, and handle the non-linear nature of the binary response variable.
+By balancing methodological rigor with practical transparency, this model ensures that the analysis remains both statistically sound and actionable for stakeholders.
+
+### Model Assumption
+
+This Bayesian model relies on several assumptions that ensure the validity of its predictions and the transparency of its coefficients.
+While these assumptions are generally less restrictive than those for simpler models like linear regression, they remain critical to the robustness of the results.
+The key assumptions are listed below:
+
+1. Binary Response Variable
+The model assumes that the response variable ($y$) is binary, taking values of either 1 (approval) or 0 (refused).
+This assumption aligns with the nature of the data, as application decisions are dichotomous outcomes.
+Any deviations from binary coding would invalidate the logistic framework and require alternative modeling approaches.
+
+2. Independence of Observations:
+The model assumes that all observations are independent of one another.
+This is reasonable for application data, as each decision is typically made independently by the committee.
+However, if clustering or dependence exists (e.g., decisions within the same planning district are correlated), the model may need to incorporate hierarchical or random effects to address these dependencies.
+
+3. No Perfect Multicollinearity
+Logistic regression assumes that the predictors are not perfectly correlated, as this would prevent the model from estimating unique coefficients.
+For example, if a predictor is a linear combination of other predictors, the model would fail to converge.
+In this analysis, categorical variables of application type and planning district are appropriately encoded to avoid such issues.
+
+4. Proper Model Specification
+The model assumes that all relevant predictors are included and correctly specified.
+Omitting important variables or including irrelevant ones could lead to biased estimates or reduced transparency.
+For example, excluding interaction terms when they are theoretically justified might result in incomplete understanding of the predictors' effects.
+
+### Model Limitation
 

### 2024-11-26 paper/paper.qmd
Upload model limitations

@@
 
-### Model Limitation
+### Potential Limitation and Addressing Violation
+
+Violations of the above assumptions can impact this model’s validity while severl solution are listed below:
+
+1. Non-linearity:
+Adding polynomial terms or splines could address non-linear relationships if diagnostic checks suggest deviations.
+
+2. Dependence:
+If observations are not independent (e.g. decisions clustered by region), hierarchical models or random effects may be introduced.
+
+3. Multicollinearity:
+Variance inflation factors or correlation matrices can help identify multicollinearity issues, which can be resolved by removing or combining correlated predictors.
+
+By adhering to these assumptions and addressing potential violations, the Bayesian logistic regression model provides a robust and effective framework for analyzing the factors influencing application outcomes.
 

### 2024-11-26 scripts/00-simulate_data.R
Revise variables

@@ -33,10 +33,6 @@ application_type <- c("MV", "CO")
 # Decision types
 decision <- c(1, 0)
 
-# Set a date range
-start_date <- as.Date("2011-02-23")
-end_date <- as.Date("2024-10-22")
-
 # Create a dataset by randomly assigning states and parties to divisions
 simulated_data <- tibble(
   application_type = sample(
@@ -45,9 +41,9 @@ simulated_data <- tibble(
     replace = TRUE,
     prob = c(0.3, 0.7) # Rough distribution
   ),
-  date = as.Date(sample(seq(start_date, end_date, by = "day"), 500,
+  year = sample(seq(2016, 2024), 500,
     replace = TRUE
-  )),
+  ),
   planning_district = sample(
     planning_district,
     size = 500,

### 2024-11-26 scripts/01-test_simulated_data.R

@@ -26,7 +26,7 @@ simulated_data <- read_parquet("data/00-simulated_data/simulated_data.parquet")
 #### Test data ####
 
 # Check if there are any missing values in the dataset
-if (all(!is.na(simulated_data))) {
+if (all(!is.na(analysis_data))) {
   message("Test Passed: The dataset contains no missing values.")
 } else {
   stop("Test Failed: The dataset contains missing values.")
@@ -44,19 +44,29 @@ check_required_columns <- function(data, required_columns) {
 
 # Check if each column has the correct data type
 check_column_types <- function(data) {
-  if (!is.Date(as.Date(data$date))) stop("Error: 'date' column must be in Date format.")
-  if (!is.character(data$application_type)) stop("Error: 'application_type' column must be character.")
-  if (!is.character(data$planning_district)) stop("Error: 'planning_district' column must be character.")
-  if (!is.numeric(data$decision)) stop("Error: 'decision' column must be character.")
+  if (!is.numeric(data$year)) {
+    stop("Error: 'year' column must be numbers.")
+  }
+  if (!is.character(data$application_type)) {
+    stop("Error: 'application_type' column must be character.")
+  }
+  if (!is.character(data$planning_district)) {
+    stop("Error: 'planning_district' column must be character.")
+  }
+  if (!is.numeric(data$decision)) {
+    stop("Error: 'decision' column must be numbers.")
+  }
   message("Check passed: All columns have the correct format.")
 }
 
-# Test if all 'date' values are between Feb 23, 2011, and Oct 22, 2024
-check_date_range <- function(data) {
-  if (!all(as.Date(data$date) >= as.Date("2011-02-23") & as.Date(data$date) <= as.Date("2024-10-22"))) {
-    stop("Error: 'date' column contains values outside the valid range (2011-02-23 to 20234-10-22).")
+# Test if all 'year' values are between 2016, and 2024
+check_year_range <- function(data) {
+  if (!all(data$year >= 2016 &
+             data$year <= 2024)) {
+    stop("Error: 'year' column contains values outside
+         the valid range (2016 to 2024).")
   } else {
-    message("Check passed: All 'date' values are within the proper range.")
+    message("Check passed: All 'year' values are within the proper range.")
   }
 }
 
@@ -72,7 +82,8 @@ check_application_types <- function(data) {
 
 # Test if 'planning_district' values are within the district list
 check_planning_districts <- function(data) {
[…]

### 2024-11-26 scripts/03-clean_data.R

@@ -26,16 +26,21 @@ raw_data$IN_DATE <- as.Date(raw_data$IN_DATE, format = "%Y-%m-%d")
 cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
-  select(application_type, in_date, planning_district, c_of_a_descision) |>
   filter(grepl("approved|refused|approval", c_of_a_descision,
-               ignore.case = TRUE)) |>
-  mutate(c_of_a_descision = ifelse(
-    grepl("approved|approval", c_of_a_descision, ignore.case = TRUE),
-    1, 0
+    ignore.case = TRUE
   )) |>
-  rename(date = in_date,
-    decision = c_of_a_descision
+  mutate(
+    c_of_a_descision = ifelse(grepl("approved|approval",
+      c_of_a_descision,
+      ignore.case = TRUE
+    ),
+    1, 0
+    ),
+    year = year(in_date)
   ) |>
+  filter(year >= 2016) |>
+  rename(decision = c_of_a_descision) |>
+  select(application_type, year, planning_district, decision) |>
   tidyr::drop_na()
 
 #### Save data ####

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -43,19 +43,29 @@ check_required_columns <- function(data, required_columns) {
 
 # Check if each column has the correct data type
 check_column_types <- function(data) {
-  if (!is.Date(as.Date(data$date))) stop("Error: 'date' column must be in Date format.")
-  if (!is.character(data$application_type)) stop("Error: 'application_type' column must be character.")
-  if (!is.character(data$planning_district)) stop("Error: 'planning_district' column must be character.")
-  if (!is.numeric(data$decision)) stop("Error: 'decision' column must be character.")
+  if (!is.numeric(data$year)) {
+    stop("Error: 'year' column must be numbers.")
+  }
+  if (!is.character(data$application_type)) {
+    stop("Error: 'application_type' column must be character.")
+  }
+  if (!is.character(data$planning_district)) {
+    stop("Error: 'planning_district' column must be character.")
+  }
+  if (!is.numeric(data$decision)) {
+    stop("Error: 'decision' column must be numbers.")
+  }
   message("Check passed: All columns have the correct format.")
 }
 
-# Test if all 'date' values are between Feb 23, 2011, and Oct 22, 2024
-check_date_range <- function(data) {
-  if (!all(as.Date(data$date) >= as.Date("2011-02-23") & as.Date(data$date) <= as.Date("2024-10-22"))) {
-    stop("Error: 'date' column contains values outside the valid range (2011-02-23 to 20234-10-22).")
+# Test if all 'year' values are between 2016, and 2024
+check_year_range <- function(data) {
+  if (!all(data$year >= 2016 &
+    data$year <= 2024)) {
+    stop("Error: 'year' column contains values outside
+         the valid range (2016 to 2024).")
   } else {
-    message("Check passed: All 'date' values are within the proper range.")
+    message("Check passed: All 'year' values are within the proper range.")
   }
 }
 
@@ -71,8 +81,10 @@ check_application_types <- function(data) {
 
 # Test if 'planning_district' values are within the district list
 check_planning_districts <- function(data) {
-  valid_districts <- c("Etobicoke York", "North York", "Scarborough",
-                       "Toronto East York")
+  valid_districts <- c(
+    "Etobicoke York", "North York", "Scarborough",
+    "Toronto East York"
+  )
   if (!all(data$planning_district %in% valid_districts)) {
     stop("Error: Invalid 'planning_district' values found.")
   } else {
@@ -91,10 +103,13 @@ check_decision_values <- function(data) {
 }
 
 # Run all checks
[…]

### 2024-11-26 scripts/06-model_data.R

@@ -23,13 +23,14 @@ analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 analysis_data <- analysis_data %>%
   mutate(
     application_type = as.factor(application_type),
+    year = as.factor(year),
     planning_district = as.factor(planning_district)
   )
 
 #### Model data ####
 first_model <- 
   stan_glm(
-    formula = decision ~ application_type + date + planning_district,
+    formula = decision ~ application_type + year + planning_district,
     data = analysis_data,
     family = binomial(link = "logit"), # Logistic regression
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-11-26 paper/paper.qmd
Upload reference and readme

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions"
+subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications leveraging Bayesian Logistic Regression"
+author: Angel Xu
+thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
 date: today
@@
 format: pdf
+toc: true
 number-sections: true
@@
 library(tidyverse)
-library(palmerpenguins)
+library(opendatatoronto)
+library(modelsummary)
+
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 ```
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+Our data [@scommittee_application]....
+
+The analyses presented in this paper were conducted using R programming language [@citeR].
+The `tidyverse` packages [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow] were used in the process of data simulation, testing beforehand.
+After the original raw data was downloaded by using `tidyverse` package [@citetidyverse], and `dplyr` package [@citedplyr], data cleaning process was done by using `tidyverse` package [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow].
+The code style is corrected by `here` package [@citehere], `lintr` package [@citelintr], and `styler` package [@citestyler].
+Then, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
+Graphs were made with `ggplot2` package [@citeggplot].
+The model results are then presented by `modelsummary` [@citemodelsummary] package.
+
+Following @tellingstories, we consider...
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+outcome variable: decision
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+@fig-decision shows the distribution of application decisions.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-decision
+#| fig-cap: Distribution of Application Decisions
 #| echo: false
+#| warning: false
+#| message: false
+
+#### Workspace Setup####
+library(ggplot2)
+
+#### Plot the distribution of `decision` ####
+ggplot(analysis_data, aes(x = factor(decision))) +
+  geom_bar(fill = "lightblue", color = "black") +
+  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  labs(
+    x = "Decision (0 = Refused, 1 = Approved)",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    axis.title.x = element_text(size = 14),
+    axis.title.y = element_text(size = 14),
+    axis.text = element_text(size = 12)
+  )
 
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
 ```
@@
 
+### Application type
+
 Add graphs, tables and text.
@@
 
+@fig-application-types presents the distribution of application types.
+
+```{r}
+#| label: fig-application-types
+#| fig-cap: Distribution of Application Types
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot the distribution of `application_type` ####
+ggplot(analysis_data, aes(x = factor(application_type))) +
+  geom_bar(fill = "lightgreen", color = "black") +
+  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  labs(
+    x = "Application Type",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    axis.title.x = element_text(size = 14),
+    axis.title.y = element_text(size = 14),
+    axis.text = element_text(size = 12)
+  )
+
+```
+
+### Year
+
+@fig-year presents the distribution of application years
+
+```{r}
+#| label: fig-year
+#| fig-cap: Distribution of Applications Years
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot the distribution of `year` ####
+ggplot(analysis_data, aes(x = factor(year))) +
+  geom_bar(fill = "skyblue", color = "black") +
+  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  labs(
+    x = "Year",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    axis.title.x = element_text(size = 14),
+    axis.title.y = element_text(size = 14),
+    axis.text = element_text(size = 12)
+  )
+
+```
+
+### Planning District
+
+@fig-district shows the distribution of planning district.
+
+```{r}
+#| label: fig-district
+#| fig-cap: Distribution of Planning District
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot the distribution of `planning_district` ####
+ggplot(analysis_data, aes(x = factor(planning_district))) +
+  geom_bar(fill = "coral", color = "black") +
+  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  labs(
+    x = "Planning District",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    axis.title.x = element_text(size = 14),
+    axis.title.y = element_text(size = 14),
+    axis.text = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate district labels for readability
+  )
+
+```
+
 # Model {#sec-model}
 
-In this paper, a Bayesian logistic model will be utilized to understand how factors such as application type, submission date, and planning district influence the decisions of the application committee.
+In this paper, a Bayesian logistic model will be utilized to understand how factors such as application type, submission year, and planning district influence the decisions of the application committee.
 This model estimates the probability of an application being approved (P(Approval)) based on the given predictors.
@@
 \begin{equation}
-\log\left(\frac{p_i}{1 - p_i}\right) = \beta_0 + \beta_1 \times \text{application type} + \beta_2 \times \text{submission date} + \beta_3 \times \text{planning district}
+\log\left(\frac{p_i}{1 - p_i}\right) = \beta_0 + \beta_1 \times \text{application type} + \beta_2 \times \text{year} + \beta_3 \times \text{planning district}
 \end{equation}
@@
 - $\beta_1$: Coefficients for application type, dummy-coded with CO as the reference category.
-- $\beta_2$: Coefficient for date, measured as the year of application to capture temporal trends.
+- $\beta_2$: Coefficient for year, measured as the year of application to capture temporal trends.
 - $\beta_3$: Coefficients for planning district, dummy-coded with North York as the reference category.

### 2024-11-28 paper/paper.qmd
Upload Data part

@@
 library(modelsummary)
+library(knitr)
 
@@
 
-Our data [@scommittee_application]....
+Our data [@committee_application] is extracted from opendatatoronto [@opendatatoronto].
+This dataset focuses on applications submitted to the City of Toronto's Committee of Adjustment, a governing body responsible for reviewing minor variance and consent applications related to land use and development.
+These applications are submitted by property owners and developers seeking permission to deviate from existing zoning regulations or to subdivide land.
+The Committee’s decisions—whether to approve or refuse an application—have significant implications for urban development, community planning, and city growth.
 
-The analyses presented in this paper were conducted using R programming language [@citeR].
+The Committee of Adjustment plays a key role in Toronto’s urban planning ecosystem.
+It ensures that deviations from zoning bylaws align with the city’s broader development objectives, maintaining a balance between flexibility for property owners and adherence to long-term planning goals.
+The decisions made by the committee are influenced by various factors, including the type of application, geographic considerations (e.g.
+planning districts), and the timing of submission.
+Understanding these decisions provides insight into urban planning dynamics and the regulatory challenges faced by a growing city like Toronto.
+
+The analysis presented in this paper were conducted using R programming language [@citeR].
 The `tidyverse` packages [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow] were used in the process of data simulation, testing beforehand.
@@
 Graphs were made with `ggplot2` package [@citeggplot].
+Tables were constructed with `knitr` package [@citeknitr].
 The model results are then presented by `modelsummary` [@citemodelsummary] package.
 
-Following @tellingstories, we consider...
+Following @tellingstories, this paper uses Bayesian Logistic Model to explore factors that affect the application results.
 
-Overview text
+Although several similar datasets with overlapping themes were considered but ultimately not used due to limitations in scope, structure, or relevance.
+For instance, the Toronto Application Information Centre Dataset provides information on broader planning applications like rezoning requests and site plan approvals, but its complexity and inclusion of unrelated application types make it unsuitable for isolating patterns related to minor variances and consents.
+The Ontario Municipal Board (OMB) Decisions Dataset includes appeal outcomes, but its focus on contested cases introduces bias and omits the routine applications central to this study.
+
+Consequently, focusing on the Committee of Adjustment Applications dataset [@committee_application] stands out for its granularity and relevance, enabling a focused analysis of approval trends and their influencing factors, which directly aligns with the study's goals.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset captures information on applications submitted to the City of Toronto’s Committee of Adjustment, providing a structured record of decisions made on zoning variances and consents.
+The dataset’s structure results from Toronto’s well-defined application process.
+When property owners or developers submit applications to the Committee of Adjustment, they provide detailed information about their project, including its location, purpose, and the specific variances or consents requested.
+These details are reviewed by committee staff and entered into a digital database, ensuring consistency and standardization.
 
-## Outcome variables
+Applications to the Committee of Adjustment are typically submitted by property owners, developers, or their authorized agents (e.g., architects or planners).
+Eligibility to submit an application requires ownership or legal rights to the property in question.
+For some application types, such as consents, additional prerequisites include demonstrating compliance with existing zoning bylaws or providing documentation showing that the proposed subdivision aligns with city planning objectives.
+Minor variance applications, on the other hand, require applicants to show that the requested adjustments are “minor” in nature, do not conflict with broader zoning intentions, and maintain compatibility with the surrounding area.
+As a result, individuals submitting these applications are often professionals with expertise in urban planning, real estate, or development, or property owners working in consultation with such professionals.
 
-outcome variable: decision
+The data collection process begins with property owners or developers submitting applications through a centralized system.
+Each application includes essential details, which are systematically verified and recorded by administrative staff.
+Following public hearings, committee decisions (approval or refusal) are added to the database, completing the data entry process.
+The dataset is then maintained and periodically published by the City of Toronto, ensuring it remains accessible and accurate for public and analytical use.
 
-@fig-decision shows the distribution of application decisions.
+The dataset thus provides a structured lens through which to examine urban planning decisions.
+By distilling real-world phenomena—such as zoning adjustments and land use approvals—into discrete variables, the dataset enables systematic analysis of trends and relationships within Toronto’s regulatory framework.
+This structured approach ensures that the dataset is both practical for analysis and representative of the nuanced processes underlying urban development.
+
+## Data Cleaning
+
+To ensure the dataset was suitable for analysis, several high-level cleaning steps were performed.
+First, unused decision outcomes were filtered out, retaining only applications that were explicitly approved or refused.
+The `decision` variable was then converted into a binary format, with `1` representing approvals and `0` representing refusals, simplifying its use in modeling.
+
+A new variable, `year`, was constructed by extracting the year from the `in_date` field.
+This allowed for a clearer examination of temporal trends while maintaining the granularity of the original data.
+Applications from very early years with minimal counts were excluded to avoid bias and ensure a representative analysis of contemporary decision-making trends.
+
+Additionally, key variables are selected out in cleaned dataset for convenience. More details will be explored in appendix.
+
+## Features
+
+The dataset provides details of applications submitted to the Committee of Adjustment, focusing on key variables:
+
+- Decision: Indicates whether an application was approved or refused.
+- Application Type: Categorizes the application as either a minor variance or a consent.
+- Year: The year when the application was submitted.
+- Planning District: The geographic area within Toronto where the application originated.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-preview
+#| tbl-cap: "Preview of Key Variables"
+
+# Show first six rows of the cleaned dataset
+analysis_data |>
+  head() |>
+  kable(booktabs = TRUE, align = "c")
+```
+
+The first six rows of the cleaned dataset, displayed in @tbl-preview, offer a clear preview of these key variables.
+
+These variables allow for an in-depth analysis of the factors influencing application decision results and provide a quantitative basis for evaluating trends in urban planning decisions.
+Specific and further discussion about these variables are included in @sec-data-outcome and @sec-data-predictor.
+
+## Outcome variables {#sec-data-outcome}
+
+The outcome variable, decision, captures whether an application submitted to the Committee of Adjustment was approved (represented as 1) or refused (represented as 0).
 
@@
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+@fig-decision illustrates the distribution of decision outcomes.
+Out of the total applications, 1,207 (66.5%) were approved, while 607 (33.5%) were refused.
+This demonstrates that the majority of applications submitted to the Committee of Adjustment are approved, suggesting that most requests align with zoning regulations or are deemed acceptable following the committee's review process.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
 #| warning: false
 #| message: false
+#| echo: false
+#| label: tbl-decision
+#| tbl-cap: "Summary Statistics of Decisions"
 
+#### Generate summary statistics for `decision` ####
+decision_summary <- analysis_data %>%
+  group_by(decision) %>%
+  summarize(
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add '%' symbol
+  ) %>%
+  mutate(Decision = ifelse(decision == 1, "Approved (1)", "Refused (0)")) %>%
+  select(Decision, Percentage)
+
+#### Create a kable table ####
+decision_summary %>%
+  kable(
+    format = "html", 
+    col.names = c("Decision", "Percentage")
+  )
 ```
 
-Talk way more about it.
+@tbl-decision provides a clear picture of the approval and refusal rates within the dataset, highlighting a relatively high likelihood of approval for submitted applications.
 
-## Predictor variables
+## Predictor variables {#sec-data-predictor}
 
@@
 
-Add graphs, tables and text.
+The application type variable categorizes each application submitted to the Committee of Adjustment into two types:
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+Consent (CO): Requests related to the subdivision of land into separate parcels, often for new developments or property sales.
 
-@fig-application-types presents the distribution of application types.
+Minor Variance (MV): Applications seeking small adjustments to zoning regulations, such as building setbacks, height limits, or lot coverage.
+
+These categories represent distinct regulatory processes, with Consents generally requiring more extensive evaluation compared to Minor Variances due to their implications on land use.
 
@@
 
+@fig-application-types presents the distribution of applications by type.
+Minor Variances account for the majority of applications which reach to 1371 (75.6%), reflecting the routine nature of these requests in urban development.
+In contrast, Consents are less frequent which only have 443 counts (24.4%), likely due to their more specific requirements and potential complexity.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-application-type
+#| tbl-cap: "Summary Statistics of Application Types"
+
+#### Summary Statistics for Application Type ####
+application_summary <- analysis_data %>%
+  group_by(application_type) %>%
+  summarize(
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add %
+  ) %>%
+  mutate(Application_Type = ifelse(application_type == "MV", "Minor Variance (MV)", "Consent (CO)")) %>%
+  select(Application_Type, Percentage)
+
+#### Create a kable table ####
+application_summary %>%
+  kable(
+    format = "html", 
+    col.names = c("Application Type", "Percentage")
+  )
+```
+
+@tbl-application-type emphasize the dominance of Minor Variances in the dataset, providing context for their role in urban development.
+
+The relationship between application_type and the outcome variable decision will be explored further to determine if approval rates differ between Minor Variances and Consents.
+Preliminary trends suggest that Minor Variances may have higher approval rates due to their less complex nature.
+Additional analyses may examine whether these patterns vary by planning district or year.
+
 ### Year
 
-@fig-year presents the distribution of application years
+The year variable represents the year of submission for each application in the dataset.
+It provides insight into temporal trends in applications submitted to the Committee of Adjustment.
+This variable is instrumental in identifying changes in the volume of applications over time, which may reflect policy shifts, urban development trends, or external factors like economic conditions.
 
@@
 
+@fig-year clarifies the distribution of applications from 2016 to 2024, with notable fluctuations in annual counts.
+Fewer applications were submitted during 2016-2019, with counts declining steadily from 198 applications in 2016 to a low of 93 in 2019.
+This period may reflect a stabilization phase or lower activity in urban development processes.
+Afterwards, application counts increased significantly, peaking in 2023 with 378 applications, followed by a slight decline to 249 in 2024.
+The rise in 2023 may be due to heightened development activity or easing of restrictions from prior external events, such as the COVID-19 pandemic.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-year
+#| tbl-cap: "Summary Statistics of Years"
+
+#### Summary Statistics for Year ####
+year_summary <- analysis_data %>%
+  group_by(year) %>%
+  summarize(
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
+  ) %>%
+  arrange(year)
+
+#### Create a kable table ####
+year_summary %>%
+  kable(
+    format = "html", 
+    col.names = c("Years", "Percentage")
+  )
+```
+
+@tbl-year states how each year contributes to the total volume of applications.
+
+The year variable is likely related to the decision outcome and application type.
+As approval rates may have varied over time due to changing city planning policies or committee priorities.
+Besides, temporal trends may reveal shifts in the proportion of Minor Variances (MV) versus Consents (CO), reflecting evolving development demands.
+
 ### Planning District
 
-@fig-district shows the distribution of planning district.
+The planning district variable identifies the geographic district within Toronto where an application was submitted.
+The city is divided into four planning districts, each with its own development characteristics and zoning regulations.
+These districts are crucial for understanding spatial trends in application submissions and outcomes, as local policies and priorities can influence decision-making processes.
 
@@
 
+@fig-district displays the distribution of applications across four planning districts.
+We can observe that North York has the highest number of applications, accounting for a significant proportion of the dataset.
+This could reflect the area's high density of development projects or active property owners seeking zoning adjustments.
+Scarborough has the fewest applications, possibly due to fewer development activities or stricter zoning compliance.
+Etobicoke York and Toronto East York have moderate numbers of applications, reflecting their roles as active but less prominent districts compared to North York.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-district
+#| tbl-cap: "Summary Statistics of Planning Districts"
+
+#### Summary Statistics for Planning District ####
+district_summary <- analysis_data %>%
+  group_by(planning_district) %>%
+  summarize(
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
+  )
+
+#### Create a kable table ####
+district_summary %>%
+  kable(
+    format = "html", 
+    col.names = c("Planning Districts", "Percentage")
+  )
+```
+
+@tbl-district provide a view of how applications are distributed spatially across the city.
+
+The planning_district variable may also interacts with application type and decision.
+Approval rates may vary by district, reflecting differences in local policies or development dynamics.
+The proportion of Minor Variances (MV) versus Consents (CO) may differ across districts, indicating distinct development priorities or zoning challenges.
+
 # Model {#sec-model}
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate, and include justification for model and the variables, as well as discuss underlying assumptions, potential limitations, software used to implement the model, and evidence of model validation and checking.
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+Here we briefly describe the Bayesian analysis model used to investigate, and include justification for model and the variables, as well as discuss underlying assumptions, potential limitations, and evidence of model validation and checking.
 
@@
 This approach allows for robust inference, offering a clear quantification of parameter uncertainty and supporting more nuanced interpretations of the predictors' effects.
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
@@
 
-### Model Implementation
-
 # Results
@@
 
-# Additional data details
+# Additional data cleaning details
+
+The dataset underwent a structured cleaning process to ensure its suitability for analysis, focusing on preparing the data while minimizing bias and enhancing interpretability.
+Below is a detailed description of the key cleaning steps:
+
+1. **Filtering Unused Decision Outcomes**
+
+The dataset originally included a range of decision outcomes beyond approvals and refusals, such as applications that were deferred or withdrawn.
+These were excluded to focus solely on cases where the Committee of Adjustment made a definitive decision.
+This exclusion is done by filtering decisions with "approve", "approval", and "refused" words in it while conditionally approved are all default as approval to enlarge the data size.
+This ensured a binary structure for the decision variable, aligning with the study’s objectives.
+
+2. **Conversion of decision to Binary Format**
+
+The decision variable was converted into a binary format for consistency and ease of analysis:
+
+1: Representing approved applications.
+0: Representing refused applications. This transformation provided a clear and standardized representation of application outcomes.
+
+3. **Creation of the year Variable**
+
+A new variable, year, was created by extracting the year component from the date field.
+This allowed for an analysis of temporal trends without losing the temporal resolution of the data.
+The original date variable remained intact for potential use in supplementary analyses.
+
+4. **Filtering Applications from Earlier Years**
+Applications before 2016 were removed from the dataset due to their low counts.
+Retaining these outliers could lead to biased results, particularly in trend analyses or modeling.
+By excluding these years, the dataset better represents the more recent and consistent patterns of decision-making.
+
+5. **Removing Missing Values**
+Records with missing values in the chosen variables (decision, application_type, year, and planning_district) were removed.
 

### 2024-11-28 scripts/06-model_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Bayesian Logisitic Regression Model
+# Author: Angel Xu
+# Date: 23 November 2024
+# Contact: anjojoo.xu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: run 01-dowload_data.R and 03-clean_data.R to get the dataset
+# Any other information needed? None
 
 
 #### Workspace setup ####

### 2024-11-28 paper/paper.qmd
Upload model

@@
 Then, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
-Graphs were made with `ggplot2` package [@citeggplot].
+Model validation utilizes `Metrics` package [@citeMetrics] Graphs were made with `ggplot2` package [@citeggplot].
 Tables were constructed with `knitr` package [@citeknitr].
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm and `Metrics` package of @citeMetrics.
+`rstanarm`@rstanarm is used to set up the model, and `Metrics` @citeMetrics is used to calculate RMSE.
 
@@
 
-### Model justification
+## Model justification
 
@@
 
-### Model Assumption
+## Model Assumption
 
@@
 
-### Potential Limitation and Addressing Violation
+## Potential Limitation and Addressing Violation
 
@@
 
+## Model Validation
+
+To ensure the robustness and reliability of the Bayesian logistic regression model, a series of validation and diagnostic steps were conducted.
+Detailed plots and results, including posterior predictive checks (PPC), posterior vs.
+prior comparisons, and convergence diagnostics, are provided in the appendix for further examination.
+Below, the focus is on other key validation methods:
+
+1. **Out-of-Sample Testing**
+
+The dataset was split into a training set (80%) and a test set (20%) to evaluate the model's predictive performance on unseen data.
+The model was trained on the training set, and predictions were made on the test set.
+Classification accuracy was calculated as the proportion of correctly classified outcomes in the test set.
+This approach assesses the model's generalizability.
+
+2. **Predictive Accuracy (RMSE)**
+
+The Root Mean Square (RMSE) typically measures the deviation between predicted values and observed values in the same metric.
+In the case of a logistic regression model predicting probabilities ($p_i$), the observed outcomes ($y_i$) are binary, which means they aren’t directly comparable in the usual sense of residuals as continuous values.
+However, RMSE can still be used as a measure of predictive performance for logistic regression by interpreting as the observed realization of the probabilities.
+
+Logistic regression models the probability of an event occurring, $\hat{p_i}=P(y_i=1)$ while The observed binary outcome (y_i) is an approximation of whether the event occurred (1) or not (0).
+The squared error between the predicted probability and the binary outcome measures how well the model assigns probabilities close to the actual outcome.
+Thus, the equation of the Squared Error is :
+
+$$Squared \: Error = (y_i-\hat{p_{i}})^2$$
+
+RMSE aggregates this squared error across all observations and provides a measure of how well the model predicts probabilities in line with the observed binary outcomes.
+Consequently, the RMSE equation is:
+
+$$RMSE = \sqrt{\frac{1}{n}\sum_{i=1}^{n}(y_{i}-\hat{p_{i}})^2}$$
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+#### Workspace setup ####
+library(Metrics) # For RMSE calculation
+library(rstanarm)
+
+#### Read the model
+model <-
+  readRDS(file = here::here("models/first_model.rds"))
+
+#### Generate predictions ####
+# Predicted probabilities
+predicted_probabilities <- posterior_predict(model, draws = 1) %>% 
+  as.vector()
+
+# RMSE Calculation
+rmse_value <- rmse(analysis_data$decision, predicted_probabilities)
+
+#### Print RMSE ####
+print(paste("RMSE:", rmse_value))
+```
+
+A low RMSE indicates that the predicted probabilities are close to the observed outcomes, demonstrating good predictive accuracy.
+
+3. Sensitivity Analysis
+The sensitivity of the model was assessed by varying the priors for the coefficients to examine whether the results were robust to different prior specifications.
+The predictions remained stable across different prior choices, indicating that the model is not overly influenced by prior assumptions.
+
+4. Predictive Power (AUC-ROC)
+The Area Under the Receiver Operating Characteristic Curve (AUC-ROC) was computed to evaluate the model's ability to discriminate between approved and refused applications.
+A high AUC-ROC value (>0.8) confirmed the model's strong discriminatory performance.
+
 # Results
@@
 #| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Model results of Application Approval Probability"
 #| warning: false
-#| message: false
 
-library(rstanarm)
+modelsummary::modelsummary(
+  list(
+    "Model" = model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
+The result highlights how the predictors influence the likelihood of application approval while controlling for other variables.
+The coefficients, presented as log-odds, provide insights into both the direction and magnitude of each predictor’s impact on the approval decision.
+
+The intercept estimate of -0.06 represents the baseline log-odds of approval when all predictors are at their reference levels.
+This corresponds to a baseline probability close to 0.5, indicating nearly equal odds of approval or refusal in the absence of other influencing factors.
+This serves as a neutral starting point for understanding how the predictors shift the likelihood of approval.
+
+The variable application_typeMV, with a coefficient of -0.20, suggests that applications for minor variances (MV) are less likely to be approved compared to consents (CO), which act as the reference category.
+In terms of odds, the transition from a consent to a minor variance reduces the odds of approval by approximately 18% ($e^{-0.20}\approx 0.82$).
+This finding aligns with expectations, as minor variances may face stricter scrutiny due to their potentially greater impact on local zoning adjustments.
+
+The year variables reveal notable temporal trends in approval likelihood, using 2016 as the reference year.
+Applications submitted in 2017 had significantly higher odds of approval, with a 109% increase in odds compared to 2016 ($e^{0.74}\approx 2.09$).
+This upward trend continued into 2018, which saw the highest approval odds, with a 161% increase over 2016 ($e^{0.96}\approx 2.61$).
+In 2019, the odds of approval remained positive but were less pronounced, with a 25% increase over 2016 ($e^{0.22}\approx 1.25$).
+By 2020, the odds rose again, with a 75% increase compared to the baseline year ($e^{0.56}\approx 1.75$).
+These results suggest that approval rates have generally improved over time, potentially reflecting shifts in policy, procedural efficiency, or external factors affecting decision-making.
+
+The planning district coefficients provide insight into geographic variability in approval rates, highlighting differences in decision-making across Toronto’s planning regions.
+North York leads in both application volume and approval odds, reflecting its role as a development hub.
+Scarborough faces lower approval odds, suggesting distinct planning challenges or fewer adjustments being approved.
+These geographic differences likely reflect localized development pressures, zoning considerations, or district-specific challenges that influence decision-making.
+
+In summary, the model reveals key patterns in how application type and submission year influence the likelihood of approval.
+Minor variance applications face slightly lower odds of approval compared to consents, while temporal trends indicate an overall increase in approval rates from 2016 to 2020.
+These findings provide a foundation for understanding the dynamics of approval decisions and highlight areas for further exploration, such as geographic variability or interactions between predictors.
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# #### Extract coefficients and statistics ####
+# # Extract summary of the model
+# model_summary <- summary(model)$coefficients
+# 
+# # Convert to a data frame
+# model_summary_df <- as.data.frame(model_summary)
+# 
+# # Add row names (predictor names) as a column
+# model_summary_df <- model_summary_df %>%
+#   rownames_to_column(var = "term") %>% # Convert row names to a column
+#   rename(
+#     Estimate = Estimate, # Mean estimate (log-odds)
+#     StdError = `Std. Error` # Rename standard error for clarity
+#   ) %>%
+#   mutate(
+#     conf.low = Estimate - 1.96 * StdError, # Confidence interval lower bound
+#     conf.high = Estimate + 1.96 * StdError # Confidence interval upper bound
+#   ) %>%
+#   filter(term != "(Intercept)") # Exclude intercept for clarity
+# 
+# #### Create the coefficient plot ####
+# ggplot(model_summary_df, aes(x = reorder(term, Estimate), y = Estimate)) +
+#   geom_point(size = 3, color = "blue") +
+#   geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "black") +
+#   coord_flip() +
+#   labs(
+#     title = "Estimated Coefficients with 95% Confidence Intervals",
+#     x = "Predictor",
+#     y = "Coefficient (Log-Odds)"
+#   ) +
+#   theme_minimal(base_size = 14) +
+#   theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 
-pp_check(first_model) +
+pp_check(model) +
   theme_classic() +
@@
 
-posterior_vs_prior(first_model) +
+posterior_vs_prior(model) +
   theme_minimal() +
@@
 
-plot(first_model, "trace")
+plot(model, "trace")
 
-plot(first_model, "rhat")
+plot(model, "rhat")
 ```

### 2024-11-28 paper/paper.qmd
Upload Model Details

@@
 
-## First discussion point {#sec-first-point}
+## Overview
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+This paper investigates the factors influencing application approvals submitted to the City of Toronto's Committee of Adjustment, using a Bayesian logistic regression model.
+By analyzing predictors such as application type, submission year, and planning district, the study provides insights into the dynamics of urban planning decisions.
+The model captures temporal trends, spatial variability, and procedural differences, offering a structured understanding of approval outcomes.
+Visualizations, including coefficient plots and effect plots, complement the analysis, helping to interpret the relationships between predictors and approval likelihood.
 
-## Second discussion point
+## Temporal Trends in Approval Rates
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+One key finding is the temporal trend in approval rates.
+The analysis reveals that approval odds have increased steadily over time, peaking in 2018.
+This trend may reflect evolving planning policies, greater efficiency in committee processes, or a shift in zoning priorities.
+The data suggests that the Committee of Adjustment has become more permissive in granting approvals, which could indicate a growing alignment between applications and city planning objectives.
+This result highlights how administrative processes adapt over time to urban development needs.
 
-## Third discussion point
+## Geographic Disparities in Approval Decisions
 
-## Weaknesses and next steps
+Spatial variability plays a significant role in application approvals.
+The findings show that North York consistently exhibits the highest approval rates, likely due to its prominence as a development hub with infrastructure supporting large-scale projects.
+Conversely, Scarborough has the lowest odds of approval, suggesting stricter zoning enforcement or less alignment with development objectives in this area.
+This geographic disparity underscores the influence of local contexts and priorities on urban planning decisions, revealing how different districts balance growth and zoning compliance.
 
-Weaknesses and next steps should also be included.
+## Weaknesses
+
+While this paper provides insights into the factors influencing application approvals, there exists several limitations.
+These weaknesses cover data-related issues, methodological constraints, and broader contextual considerations.
+
+1. **Data Limitations**
+The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain critical variables.
+Key contextual factors, such as community objections, project complexity, or the presence of political or economic pressures, are not captured.
+For instance, the dataset does not provide information about why certain applications are refused or approved beyond the recorded decision, limiting the ability to analyze qualitative or situational influences.
+
+2. **Methodological Constraints**
+The Bayesian logistic regression model assumes that the relationships between predictors and the outcome variable are additive and linear on the log-odds scale.
+This assumption may oversimplify real-world interactions between variables, such as how specific planning districts may interact with application types or temporal trends.
+Furthermore, the model does not fully account for potential multicollinearity among predictors, which could distort the estimated coefficients.
+While priors were carefully chosen, their influence on the results might vary with a different prior specification, introducing some degree of subjectivity into the analysis.
+
+3. **Simplified Representation of Approval Dynamics**
+The approval process involves a combination of bureaucratic procedures, stakeholder negotiations, and technical considerations that are not fully captured in this analysis.
+For example, public hearings, lobbying efforts, and input from local communities likely play a significant role in shaping decisions but are absent from the dataset.
+This simplification means that the model captures only part of the story, potentially overlooking nuanced dynamics in urban planning and governance.
+
+4. **Geographic and Temporal Generalizability**
+The analysis focuses exclusively on the City of Toronto, meaning the findings are context-specific and may not generalize to other municipalities with different urban planning frameworks, policies, or demographic pressures.
+Temporal trends in approval rates, for example, could be influenced by unique local events or policy changes that are not representative of broader patterns.
+Similarly, the observed geographic variability in approval rates might reflect Toronto-specific zoning challenges rather than universal principles.
+
+## Nest Steps
+
+Building on the findings and addressing the limitations of this paper, there are several avenues for future research and analysis that could provide a deeper understanding of the factors influencing application approvals in urban planning contexts.
+
+One critical next step is to expand the dataset by incorporating additional variables that capture contextual and qualitative factors influencing decisions.
+Data on community feedback and objections during public hearings is a good choice since it could provide insights into how local resistance or support impacts approval likelihood.
+Information on project complexity, such as the scale, type, and potential impact of proposals, could reveal patterns in decision-making.
+Besides, access to committee meeting notes or further exploration in meeting minutes would help clarify the reasoning behind specific approvals or refusals.
+Furthermore, integrating economic and political context, such as local economic indicators or changes in political leadership, would offer a broader understanding of approval trends.
+
+The current model assumes additive relationships between predictors and approval likelihood, which may oversimplify real-world dynamics.
+Future research could explore nonlinear relationships, such as whether certain predictors have diminishing or exponential effects on approval rates.
+Interaction effects between predictors, such as the interplay between planning districts and application types, could also provide more nuanced insights.
+For instance, minor variances may be treated differently in densely populated urban areas compared to suburban districts.
+
+Extending this analysis to other municipalities or regions would help determine whether the observed trends in Toronto are unique or indicative of broader urban planning practices.
+Cross-regional comparisons could identify common principles across jurisdictions, highlight local deviations driven by distinct policies or demographics, and establish benchmarks for understanding how Toronto’s process aligns with other cities.
+Such studies could enhance the generalizability of findings and provide practical lessons for urban planners in different contexts.
+
+Despite to future improvements for addressing limitations, further exploration can be continued based on the current result.
+The temporal trends identified in this paper suggest shifts in approval rates over time, but the reasons behind these changes remain unexplored.
+Future work could examine specific policy changes or regulatory updates to understand their impact on approval trends.
+Additionally, external events such as economic downturns, housing crises, or infrastructure investments could be analyzed for their influence on application volumes and decisions.
+A difference-in-differences approach could also help isolate the effects of significant policy interventions, providing a clearer picture of causal relationships.
+
+Given the importance of geographic variability in approval rates, future research could adopt spatial analysis techniques.
+Spatial regression models could account for spatial autocorrelation and clustering effects in planning districts, while network analysis could study relationships between districts, such as spillover effects of development activity.
+Additionally, analyzing proximity to key infrastructure, such as transit hubs or commercial centers, could provide insights into spatial patterns of decision-making.
+
+Finally, the findings from this paper could inform the development of predictive tools to assist urban planners and developers.
+These tools could provide probabilistic forecasts of approval likelihood based on project attributes and contextual factors.
+Additionally, they could help committees identify systemic biases or inefficiencies in their decision-making processes and aid policymakers in designing more equitable and transparent approval frameworks.
+
+In summary, future research should prioritize richer data integration, explore more sophisticated modeling techniques, and broaden the scope of analysis to other regions and temporal contexts.
+These steps would not only enhance academic understanding of urban planning decisions but also provide actionable insights for policymakers, developers, and urban planners working toward effective and equitable processes.
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
+This shows the model's ability to generate data consistent with the observed data.
+By overlaying the observed data and simulated replicated datasets  generated from the posterior predictive distribution, we assess the model's adequacy.
+The close alignment between the observed data and the majority of the simulated replicated datasets indicates that the model captures the underlying structure of the data well.
+Deviations would signal potential misfits or areas where the model assumptions may need to be revised.
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+This comparison highlights how the data influenced parameter estimates.
+Large shifts from the prior to the posterior, such as seen for certain planning districts and years, suggest strong data-driven updates.
+Conversely, parameters where the posterior closely resembles the prior indicate minimal influence from the observed data.
+This evaluation helps assess the informativeness of the priors and the robustness of the parameter estimates, ensuring the model is both data- and prior-consistent.
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+It shows he sampled values for each chain of the Markov Chain Monte Carlo (MCMC) simulation over iterations.
+The chains appear to mix well and explore the parameter space without any evident trends or stickiness, suggesting convergence.
+The consistent overlap of chains indicates that the posterior distribution has been effectively sampled, reducing the risk of bias due to poor chain mixing or non-convergence.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+It evaluates the Gelman-Rubin convergence diagnostic.
+All parameters show $\hat{R}$ values close to 1, which indicates that the between-chain and within-chain variances are nearly identical.
+This confirms that the MCMC chains have converged and are sampling from the same target distribution, validating the reliability of the posterior estimates.
 
@@
 
+## Surveys, Sampling, and Observational Data
+
 \newpage

### 2024-11-29 paper/paper.qmd
Delete unused scripts

@@
 ---
-title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions"
-subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications leveraging Bayesian Logistic Regression"
+title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)"
+subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
 author: Angel Xu
@@
 Then, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
-Model validation utilizes `Metrics` package [@citeMetrics] Graphs were made with `ggplot2` package [@citeggplot].
+Graphs were made with `ggplot2` package [@citeggplot].
 Tables were constructed with `knitr` package [@citeknitr].
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm and `Metrics` package of @citeMetrics.
-`rstanarm`@rstanarm is used to set up the model, and `Metrics` @citeMetrics is used to calculate RMSE.
+We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm. `rstanarm`@citerstanarm is used to set up the model.
 
@@
 
-2. **Predictive Accuracy (RMSE)**
-
-The Root Mean Square (RMSE) typically measures the deviation between predicted values and observed values in the same metric.
-In the case of a logistic regression model predicting probabilities ($p_i$), the observed outcomes ($y_i$) are binary, which means they aren’t directly comparable in the usual sense of residuals as continuous values.
-However, RMSE can still be used as a measure of predictive performance for logistic regression by interpreting as the observed realization of the probabilities.
-
-Logistic regression models the probability of an event occurring, $\hat{p_i}=P(y_i=1)$ while The observed binary outcome (y_i) is an approximation of whether the event occurred (1) or not (0).
-The squared error between the predicted probability and the binary outcome measures how well the model assigns probabilities close to the actual outcome.
-Thus, the equation of the Squared Error is :
-
-$$Squared \: Error = (y_i-\hat{p_{i}})^2$$
-
-RMSE aggregates this squared error across all observations and provides a measure of how well the model predicts probabilities in line with the observed binary outcomes.
-Consequently, the RMSE equation is:
-
-$$RMSE = \sqrt{\frac{1}{n}\sum_{i=1}^{n}(y_{i}-\hat{p_{i}})^2}$$
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-#### Workspace setup ####
-library(Metrics) # For RMSE calculation
-library(rstanarm)
-
-#### Read the model
-model <-
-  readRDS(file = here::here("models/first_model.rds"))
-
-#### Generate predictions ####
-# Predicted probabilities
-predicted_probabilities <- posterior_predict(model, draws = 1) %>% 
-  as.vector()
-
-# RMSE Calculation
-rmse_value <- rmse(analysis_data$decision, predicted_probabilities)
-
-#### Print RMSE ####
-print(paste("RMSE:", rmse_value))
-```
-
-A low RMSE indicates that the predicted probabilities are close to the observed outcomes, demonstrating good predictive accuracy.
+2. Confusion Metrix
 
@@
 #| eval: true
+#| warning: false
+#| message: false
+
+library(rstanarm)
+
+model <-
+  readRDS(file = here::here("models/first_model.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
 #| label: tbl-modelresults
@@
 #| warning: false
+
+model <- model
 
@@
 
+### Overview
+This appendix delves into the methodological considerations underlying the data used in this paper, focusing on the challenges and intricacies of observational data, the implications of sampling, and the potential use of surveys and simulations to enrich insights.
+
+### Methodology
+
+### Sampling Approach
+
+#### Target Population
+
+#### Sampling Frame
+
+#### Trade-offs
+
+### Idealized Methodology
+
+### Survey
+
 \newpage

### 2024-11-29 scripts/05-model_data.R

@@ -28,7 +28,7 @@ analysis_data <- analysis_data %>%
   )
 
 #### Model data ####
-first_model <- 
+first_model <-
   stan_glm(
     formula = decision ~ application_type + year + planning_district,
     data = analysis_data,
@@ -47,5 +47,5 @@ saveRDS(
 
 
 #### Lint and style the code ####
-lint(filename = here("scripts/00-simulate_data.R"))
-style_file(path = here("scripts/00-simulate_data.R"))
+lint(filename = here("scripts/05-model_data.R"))
+style_file(path = here("scripts/05-model_data.R"))

### 2024-12-01 paper/paper.qmd
Upload appendix and result

@@
 Then, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
-Graphs were made with `ggplot2` package [@citeggplot].
+Graphs were made with `ggplot2` package [@citeggplot] and `tidybayes` package [@citetidybayes].
 Tables were constructed with `knitr` package [@citeknitr].
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm. `rstanarm`@citerstanarm is used to set up the model.
+We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm.
+`rstanarm` [@citerstanarm] is used to set up the model.`tidybayes` package [@citetidybayes] is used to plot predicted probabilities in for result analysis.
 
@@
 
-The intercept estimate of -0.06 represents the baseline log-odds of approval when all predictors are at their reference levels.
+The intercept estimate of -0.06 represents the baseline (an application of CO subimited in 2016 at Etobicoke York) log-odds of approval when all predictors are at their reference levels.
 This corresponds to a baseline probability close to 0.5, indicating nearly equal odds of approval or refusal in the absence of other influencing factors.
@@
 
-The year variables reveal notable temporal trends in approval likelihood, using 2016 as the reference year.
-Applications submitted in 2017 had significantly higher odds of approval, with a 109% increase in odds compared to 2016 ($e^{0.74}\approx 2.09$).
-This upward trend continued into 2018, which saw the highest approval odds, with a 161% increase over 2016 ($e^{0.96}\approx 2.61$).
-In 2019, the odds of approval remained positive but were less pronounced, with a 25% increase over 2016 ($e^{0.22}\approx 1.25$).
-By 2020, the odds rose again, with a 75% increase compared to the baseline year ($e^{0.56}\approx 1.75$).
-These results suggest that approval rates have generally improved over time, potentially reflecting shifts in policy, procedural efficiency, or external factors affecting decision-making.
-
-The planning district coefficients provide insight into geographic variability in approval rates, highlighting differences in decision-making across Toronto’s planning regions.
-North York leads in both application volume and approval odds, reflecting its role as a development hub.
-Scarborough faces lower approval odds, suggesting distinct planning challenges or fewer adjustments being approved.
-These geographic differences likely reflect localized development pressures, zoning considerations, or district-specific challenges that influence decision-making.
+The year variables reveal significant temporal variations in approval likelihood, using 2016 as the reference year.
+These coefficients highlight how approval dynamics have shifted over time.
+For instance, in 2017, the coefficient of 0.74 indicates a 109% increase in the odds of approval compared to 2016 ($e^{0.74} \approx 2.09$), which suggests a substantial policy or procedural shift.
+In 2018, the strongest positive effect is observed, with a coefficient of 0.96, reflecting a 161% increase in odds relative to 2016 ($e^{0.96} \approx 2.61$).
+However, in 2019, the coefficient drops to 0.22, representing a smaller increase of 25% in odds ($e^{0.22} \approx 1.25$), suggesting that this period may have faced stricter reviews or procedural adjustments.
+By 2020, the coefficient rebounds to 0.56, translating to a 75% increase in odds ($e^{0.56} \approx 1.75$).
+
+For more recent years, 2021 sees a coefficient of 0.34, showing a 40% increase in odds ($e^{0.34} \approx 1.40$).
+In 2022, the coefficient is 0.28, reflecting a 32% increase ($e^{0.28} \approx 1.32$).
+In 2023, approvals surged again, with a coefficient of 0.99, representing a 170% increase in odds ($e^{0.99} \approx 2.69$).
+Finally, 2024 showed the highest increase, with a coefficient of 1.24, translating to a 245% increase in approval odds compared to the baseline year of 2016 ($e^{1.24} \approx 3.46$).
+This steady upward trend reflects systemic improvements in the approval process, possibly due to increased efficiency, evolving zoning policies, or external influences favoring development.
+
+The model captures geographic variability in approval rates across Toronto’s planning districts.
+North York, with a coefficient of 0.41, shows a 50% increase in odds compared to the baseline district ($e^{0.41} \approx 1.50$), highlighting its role as a development hub and its alignment with urban growth priorities.
+Scarborough, on the other hand, has the lowest approval likelihood, with a coefficient of 0.16, indicating a modest 17% increase in odds ($e^{0.16} \approx 1.17$), suggesting challenges related to stricter zoning rules or community resistance to changes.
+Toronto East York has a coefficient of 0.33, showing a 39% increase in odds ($e^{0.33} \approx 1.39$), reflecting its mix of heritage preservation and urban development priorities.
+The results emphasize that approval rates vary significantly across districts due to localized zoning policies, development pressures, and socio-political factors influencing decision-making.
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# #### Extract coefficients and statistics ####
-# # Extract summary of the model
-# model_summary <- summary(model)$coefficients
-# 
-# # Convert to a data frame
-# model_summary_df <- as.data.frame(model_summary)
-# 
-# # Add row names (predictor names) as a column
-# model_summary_df <- model_summary_df %>%
-#   rownames_to_column(var = "term") %>% # Convert row names to a column
-#   rename(
-#     Estimate = Estimate, # Mean estimate (log-odds)
-#     StdError = `Std. Error` # Rename standard error for clarity
-#   ) %>%
-#   mutate(
-#     conf.low = Estimate - 1.96 * StdError, # Confidence interval lower bound
-#     conf.high = Estimate + 1.96 * StdError # Confidence interval upper bound
-#   ) %>%
-#   filter(term != "(Intercept)") # Exclude intercept for clarity
-# 
-# #### Create the coefficient plot ####
-# ggplot(model_summary_df, aes(x = reorder(term, Estimate), y = Estimate)) +
-#   geom_point(size = 3, color = "blue") +
-#   geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "black") +
-#   coord_flip() +
-#   labs(
-#     title = "Estimated Coefficients with 95% Confidence Intervals",
-#     x = "Predictor",
-#     y = "Coefficient (Log-Odds)"
-#   ) +
-#   theme_minimal(base_size = 14) +
-#   theme(plot.title = element_text(hjust = 0.5))
-```
+#| label: fig-pp-type
+#| fig-cap: Predicted Probability of Approval by Application Type
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load required libraries
+library(tidybayes)
+
+# Clean analysis_data to drop unused levels
+new_data <- analysis_data %>%
+  mutate(
+    application_type = factor(application_type),
+    year = factor(year),
+    planning_district = factor(planning_district)
+  )
+
+# Generate prediction data without creating NAs
+predict_data <- expand_grid(
+  application_type = levels(new_data$application_type),
+  year = levels(new_data$year),
+  planning_district = levels(new_data$planning_district)
+)
+
+# Generate predicted probabilities
+posterior_probs <- posterior_epred(model, newdata = predict_data)
+
+# Transpose posterior_probs to match predict_data
+posterior_probs_t <- t(posterior_probs)
+
+# Check dimensions for consistency
+if (nrow(predict_data) != nrow(posterior_probs_t)) {
+  stop("Dimension mismatch between predict_data and posterior_probs.")
+}
+
+# Convert predictions to long format for boxplots
+predict_data_long <- predict_data %>%
+  mutate(row_id = row_number()) %>%
+  bind_cols(as_tibble(posterior_probs_t)) %>%
+  pivot_longer(cols = starts_with("V"), names_to = "sample", values_to = "prob")
+
+# Plot boxplot for application type
+ggplot(predict_data_long, aes(x = application_type, y = prob)) +
+  geom_boxplot() +
+  labs(
+    x = "Application Type",
+    y = "Predicted Probability of Approval"
+  ) +
+  theme_minimal()
+
+```
+
+@fig-pp-type demonstrates variability in the predicted probability of approval between consent applications (CO) and minor variances (MV).
+Consent applications exhibit a slightly higher median predicted probability of approval compared to minor variances.
+The interquartile range (IQR) for both application types is comparable, indicating a similar spread of probabilities.
+However, minor variance applications show a broader lower whisker, suggesting that certain cases have substantially lower probabilities of approval.
+This supports the interpretation that minor variances may face stricter scrutiny due to their greater potential impact on zoning.
+
+```{r}
+#| label: fig-pp-year
+#| fig-cap: Predicted Probability of Approval by Year
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generate predicted probabilities for year
+ggplot(predict_data_long, aes(x = year, y = prob)) +
+  geom_boxplot() +
+  labs(
+    x = "Year",
+    y = "Predicted Probability of Approval"
+  ) +
+  theme_minimal()
+```
+
+@fig-pp-year explores temporal trends in the predicted probability of approval from 2016 to 2024.
+The median predicted probability of approval rises steadily from 2016 to 2024.
+While 2016 exhibits the lowest probabilities, years such as 2018 and 2024 display higher medians, indicating increased approval likelihood in recent years.
+Variability within years remains relatively stable, with slight fluctuations in the spread across the interquartile range.
+This pattern aligns with findings from the regression analysis, where temporal predictors indicated a gradual improvement in approval rates over time.
+This likely reflects policy or procedural changes that have positively influenced the decision-making process.
+
+```{r}
+#| label: fig-pp-district
+#| fig-cap: Predicted Probability of Approval by Planning District
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generate predicted probabilities for planning district
+ggplot(predict_data_long, aes(x = planning_district, y = prob)) +
+  geom_boxplot() +
+  labs(
+    x = "Planning District",
+    y = "Predicted Probability of Approval"
+  ) +
+  theme_minimal()
+```
+
+@fig-pp-district evaluates geographic variability in approval probabilities across the four planning districts: Etobicoke York, North York, Scarborough, and Toronto East York.
+North York consistently shows the highest median predicted probability of approval, reaffirming its status as a development hub with favorable zoning outcomes.
+In contrast, Scarborough exhibits the lowest median and a wider spread of probabilities, with outliers indicating challenging cases.
+Toronto East York and Etobicoke York present intermediate medians, suggesting more balanced approval probabilities in these regions.
+These geographic differences highlight the influence of localized development dynamics and zoning considerations on approval outcomes.
 
@@
 #| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
+#| layout-ncol: 1
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
@@
 #| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| layout-ncol: 1
 
@@
 
-## Surveys, Sampling, and Observational Data
+## Idealized Methodology and Survey
 
 ### Overview
-This appendix delves into the methodological considerations underlying the data used in this paper, focusing on the challenges and intricacies of observational data, the implications of sampling, and the potential use of surveys and simulations to enrich insights.
-
-### Methodology
+This section outlines an idealized methodology and survey design to supplement the observational dataset used in this study.
+The objective is to enhance the analysis of urban planning decisions by incorporating a carefully designed survey targeting key stakeholders in Toronto's zoning and land-use processes.
 
@@
 #### Target Population
+The target population comprises stakeholders involved in the Committee of Adjustment's decision-making process, including:
+
+- **Property Owners and Developers**: Individuals or entities who submit applications for minor variances or consents.
+- **Urban Planners and Architects**: Professionals who prepare applications or advise clients on zoning requirements.
+- **Committee Members and Staff**: Officials responsible for evaluating applications and making decisions.
+- **Community Members**: Residents who may provide feedback during public hearings.
 
 #### Sampling Frame
+The sampling frame would be constructed using publicly available application records and contact information from municipal directories.
+This ensures representation from all planning districts and captures diverse perspectives across geographic and professional domains.
+
+#### Sample Size
+Using stratified random sampling, the sample size would aim for approximately 3000 respondents, ensuring statistical validity while considering feasibility.
+
+#### Stratified Random Sampling
+Stratified random sampling enhances the accuracy of survey estimates because it ensures proportional representation of different demographic groups.
+This minimizes the risk of sampling bias, where certain groups may be overrepresented or underrepresented, skewing the results.
+
+Hence, to achieve a representative sample, stratified random sampling will be employed.
+This method involves dividing the population into homogeneous subgroups based on key demographic characteristics, then randomly selecting samples from each stratum proportionate to its size in the population.
+The strata will include:
+
+- Application types: Categorize applicants based on the purpose of their submission. (Minor Variance and Consent)
+- Planning districts: Stratify by Toronto's four major planning districts to capture geographic diversity. (North York, Scarborough, Toronto East York, and Etobicoke York)
+- Applicant roles: Group participants by their involvement in the application process. (Individual Property Owners, Developers, and Consultants)
+
+Within these strata, respondents will be randomly selected to align with the distribution of these categories in the broader population.
+The process begins by grouping the population into strata defined by characteristics such as application type, planning district, and applicant role.
+Afterward, random samples are drawn independently from each subgroup, ensuring proportional representation.
+In cases where smaller groups are critical to the study, adjustments can be made to ensure adequate representation.
+Finally, the selected samples from each subgroup are combined to create the overall sample, ensuring that all significant categories are included.
+
+This method reduces sampling variability, enhances representation, and provides more accurate estimates compared to simple random sampling, as the unique characteristics of each subgroup are reflected in the final sample.
 
 #### Trade-offs
-
-### Idealized Methodology
+While stratified random sampling ensures proportional representation and enhances the precision of survey estimates, it is not without challenges.
+One notable trade-off is the complexity involved in defining and managing multiple strata, particularly when the strata are based on multiple dimensions such as application type, planning district, and applicant role.
+This complexity increases the time and resources required for sample design and execution.
+For instance, ensuring proportional representation across strata like **Scarborough developers** or **North York consultants** may necessitate oversampling in smaller groups, leading to additional effort in recruitment and logistical coordination.
+
+Another trade-off lies in the potential for overstratification, which can result in small sample sizes within certain strata.
+If some strata, such as **Scarborough Consent applications**, have very few members, it may be difficult to draw statistically meaningful conclusions.
+This could necessitate adjustments, such as combining smaller strata or applying post-stratification weighting, which could introduce bias or reduce interpretability.
+Additionally, the method assumes that the characteristics used to define strata are adequately known and relevant, which may not account for unobserved variables that influence the outcomes, such as the socio-political environment or project complexity.
+Despite these challenges, the benefits of stratified random sampling in terms of improved accuracy and subgroup analysis make it a robust choice for this study.
+
+### Data Validation
+Survey responses would be validated through cross-referencing with the cleaned dataset. For instance:
+
+Matching survey-reported outcomes with recorded decisions.
+Verifying consistency between applicant-reported application details and dataset attributes (e.g., application type, planning district).
+To ensure representativeness, response rates would be monitored, and additional recruitment efforts would target underrepresented groups or districts.
 
@@
 
+This survey will be conducted using Google Forms, which is an effective platform for data
+collection. The survey can be accessed by the link, [Google Form Survey](https://forms.gle/SVyeccgCSsjhQzWX7).
+
+#### Survey Structure
+
+**Title:**
+
+Understanding Urban Planning Decisions
+
+**Introduction:**
+
+I appreciate your participation in this survey, which seeks to better understand the factors influencing urban planning decisions and application outcomes in Toronto.
+Your responses are crucial for enhancing the research and contributing to insights that could inform future urban planning processes.
+
+**Please note:**
+
+- Your answers will be treated with complete confidentiality.
+- Participation in this survey is voluntary.
+- You are encouraged to provide honest and thoughtful responses.
+- The survey is estimated to take about 5 minutes to complete.
+- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
+
+Thank you for your valuable contribution! As a token of our appreciation, each participant will receive $5 upon completion of the survey.
+
+**Section 1: Application Details**
+
+What type of application did you submit?
+
+- MV (Minor Variance)
+
+- CO (Consent)
+
+- Prefer not to say
+
+Which planning district was your application associated with?
+
+- North York
+
+- Scarborough
+
+- Etobicoke York
+
+- Toronto East York
+
+- Prefer not to say
+
+Which year did you submit your application?
+
+- 2016
+
+- 2017
+
+- 2018
+
+- 2019
+
+- 2020
+
+- 2021
+
+- 2022
+
+- 2023
+
+- 2024
+
+- Other (please specify)
+
+What was the outcome of your application?
+
+- Approved
+
+- Refused
+
+- Prefer not to say
+
+- Others (please specify)
+
+**Section 2: Applicant Background**
+
+What role best describes your involvement in the application process?
+
+- Property Owner
+
+- Developer
+
+- Consultant
+
+- Prefer not to say
+
+- Others (please specify)
+
+How many years of experience do you have in urban planning or related fields?
+
+- <1 year
+
+- 1–5 years
+
+- 5–10 years
+
+- >10 years
+
+- Prefer not to say
+
+**Section 3: Process and Challenges**
+
+What was the primary purpose of your application?
+
+- Residential Adjustment
+
+- Commercial Development
+
+- Mixed-Use Development
+
+- Prefer not to say
+
+- Others (please specify)
+
+What challenges did you encounter during the application process?
+
+- Zoning Compliance
+
+- Documentation Requirements
+
+- Public Objections
+
+- Prefer not to say
+
+- Others (please specify)
+
+**Section 4: Influences on Decisions**
+
+Do you believe community feedback impacted the outcome of your application?
+
+- Yes
+
+- No
+
+Were there any external factors (e.g., economic conditions, policy changes) that influenced the decision?
+
+- Yes
+
+- No
+
+Do you have any further comments or insights about the factors that might affect the application decision?
+
+[answer box]
+
+**Section 5: Feedback and Improvements**
+
+What improvements would you recommend for the application process?
+
+- Streamlining Documentation
+
+- Clearer Guidelines
+
+- Enhanced Community Engagement
+
+- Others (please specify)
+
+Do you have any further comments or insights about the application process?
+
+d[answer box]
+
+**End Message:**
+Thank You for Completing the Survey!
+
+We sincerely appreciate your time and thoughtful responses.
+Your participation is essential in helping us gain a deeper understanding of the factors influencing urban planning decisions in Toronto.
+Your insights will contribute to research aimed at improving zoning application processes and understanding decision dynamics.
+
+If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
+
+As a token of our gratitude, each participant will receive a $5 reward shortly after completing the survey. Thank you again for your valuable contribution!
+
+### Surevey Design Consideration
+
+When designing the survey, several key factors were considered to ensure representativeness, clarity, and relevance to the study's objectives.
+The survey targets individuals involved in Toronto’s urban planning processes, including property owners, developers, and consultants.
+
+Question phrasing was tailored to align with terminology familiar to the target respondents, minimizing ambiguity.
+Efforts were also made to limit the survey's length to encourage participation while ensuring sufficient data for meaningful analysis.
+A balance was struck between closed-ended questions, which facilitate quantitative analysis, and open-ended questions, which capture nuanced qualitative insights.
+Additionally, the survey design accounted for ethical considerations, such as maintaining confidentiality and emphasizing voluntary participation.
+
+### Strength
+
+The survey design has several strengths.
+Stratified random sampling ensures that the results are representative of the population, reducing sampling bias.
+By aligning strata with key demographic and process variables, the survey facilitates detailed subgroup analysis, enhancing the depth of insights.
+The inclusion of both quantitative and qualitative questions allows for an overall understanding of the factors influencing urban planning decisions.
+Closed-ended questions enable statistical analysis, while open-ended responses provide richer context and identify factors not captured in the observational dataset.
+
+The survey also addresses gaps in the cleaned dataset, such as perceptions of community feedback, challenges during the application process, and external influences on decisions.
+These insights complement the quantitative data, offering a more holistic view of the decision-making process.
+
+### Weakness
+
+Despite its strengths, the survey has some limitations.
+Stratified random sampling requires precise knowledge of the population's composition, which may not always be accurate, particularly for less-documented strata.
+Smaller strata, such as applications from certain districts or less common applicant roles, may result in small sample sizes, limiting statistical power.
+Additionally, reliance on self-reported data introduces the potential for response bias, as participants may overstate or understate certain factors, particularly regarding the impact of community feedback or external influences.
+
+The survey's limited length, while improving completion rates, may constrain the depth of responses, especially for complex topics.
+Furthermore, the survey results are context-specific, reflecting Toronto’s unique regulatory framework and urban planning dynamics, which may limit the generalizability of findings to other cities or contexts.
+Despite these challenges, the survey is a valuable tool for supplementing the dataset and addressing research objectives.
+
 \newpage

### 2024-12-01 paper/paper.qmd
Upload Estimand

@@
 ## Estimand
+
+This study aims to estimate the factors influencing the approval decisions for urban planning applications in Toronto.
+By analyzing variables such as the type of application (minor variance or consent), the year of submission, and the planning district, this study seeks to understand how these elements shape the decision-making process.
+The primary objective is to determine the causal effects of these factors on approval outcomes, thereby offering insights into the dynamics of urban planning and regulatory processes.
 

### 2024-12-02 paper/paper.qmd
Upload abstract and introduction

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper analyzes urban planning application approvals submitted to Toronto's Committee of Adjustment from 2016 to 2024, focusing on key predictors such as application type, submission year, and planning district using Bayesian logistic regression.
+The analysis shows that consents are more likely to be approved than minor variances, with approval rates generally increasing over time, particularly from 2017 onward, though fluctuations are observed in certain years.
+Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews.
+These findings enhance understanding of urban planning decision-making processes, offering insights into temporal trends, geographic variability, and areas for equitable policy adjustments."
 format: pdf
@@
 
-Overview paragraph
-
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Urban planning decisions shape the development of cities, influencing how communities grow and adapt to changing social, economic, and environmental conditions.
+In Toronto, one of Canada’s largest and most diverse cities, the approval of planning applications directly impacts growth, housing availability, and community design.
+Despite the importance of this process, limited research has explored the specific factors that influence approval decisions, particularly for minor variances and consents.
+This lack of understanding presents challenges for creating equitable and effective policies that balance the needs of stakeholders while adhering to regulatory frameworks.
+
+This paper examines the factors influencing approval decisions for planning applications submitted to the Committee of Adjustment in Toronto between 2016 and 2024.
+By analyzing application type, year of submission, and planning district, this study identifies patterns and trends in approval rates.
+Using a statistical modeling approach, it investigates how administrative processes, policy changes, and regional differences affect these outcomes over time.
+This analysis addresses a gap in urban planning research, providing a systematic evaluation of the approval process and its determinants.
+
+The findings indicate that consents are more likely to be approved than minor variances, reflecting differences in the scrutiny applied to these application types.
+Approval rates show notable temporal trends, with the highest approval likelihood observed in 2018, where odds of approval were 161% higher compared to the baseline year of 2016.
+Significant increases are also evident in 2017 and 2020, with approval rates showing consistent improvement over time, despite some fluctuations.
+Additionally, geographic disparities are evident, with North York exhibiting the highest approval probabilities, while Scarborough demonstrates lower likelihoods, potentially due to district-specific planning challenges.
+The difference in approval rates across Toronto’s planning districts suggests localized factors influence decision-making.
+These results enhance understanding of the approval process and highlight areas where decision-making could become more consistent and equitable.
+
+The remainder of this paper is structured as follows.
+@sec-data introduces the overview of the data (@sec-data-overview), measurement (@sec-data-measurement), data cleaning process (@sec-data-cleaning), data features (@sec-data-feature), as well as explanations, descriptions, table and graph summaries of outcome (@sec-data-outcome) and predictor variables (@sec-data-predictor) of the study.
+@sec-model explains the modeling process in detail, including procedure of model set-up, model justifications, model assumptions, and potential limitations.
+Then, @sec-result highlights the prediction outcome and results by model summary and plots.
+Finally, @sec-dis includes discussions based on the results, linking it to the real world and also states some strengths and weakness of the study.
 
@@
 
-## Overview
+## Overview {#sec-data-overview}
 
@@
 
-## Measurement
+## Measurement {#sec-data-measurement}
 
@@
 
-## Data Cleaning
+## Data Cleaning {#sec-data-cleaning}
 
@@
 
-## Features
+## Features {#sec-data-feature}
 
@@
 
-2. Confusion Metrix
-
-3. Sensitivity Analysis
-The sensitivity of the model was assessed by varying the priors for the coefficients to examine whether the results were robust to different prior specifications.
-The predictions remained stable across different prior choices, indicating that the model is not overly influenced by prior assumptions.
-
-4. Predictive Power (AUC-ROC)
-The Area Under the Receiver Operating Characteristic Curve (AUC-ROC) was computed to evaluate the model's ability to discriminate between approved and refused applications.
-A high AUC-ROC value (>0.8) confirmed the model's strong discriminatory performance.
-
-# Results
+2. **Confusion Metrix**
+The confusion matrix was computed to evaluate the model's classification performance, providing measures such as accuracy, sensitivity, and specificity.
+This validation assesses the model's predictive capabilities and its performance in correctly classifying approval decisions.
+
+# Results{#sec-result}
 
@@
 
-# Discussion
+# Discussion{#sec-dis}
 

### 2024-12-02 scripts/05-model_data.R

@@ -11,12 +11,13 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 library(lintr)
 library(styler)
 library(here)
 
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 
 # Ensure categorical variables are treated as factors

### 2024-12-02 scripts/06-model_validation.R

@@ -1,0 +1,54 @@
+#### Preamble ####
+# Purpose: Bayesian Logisitic Regression Model
+# Author: Angel Xu
+# Date: 23 November 2024
+# Contact: anjojoo.xu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: run 01-dowload_data.R, 03-clean_data.R to get the dataset
+# Any other information needed? None
+
+
+#### Workspace setup ####
+set.seed(853)
+library(dplyr)
+library(arrow)
+library(lintr)
+library(styler)
+library(here)
+
+#### Read data ####
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+
+
+#### 80% of the sample size
+size <- floor(0.8 * nrow(analysis_data))
+
+# Split data into training and test sets
+train_index <- sample(seq_len(nrow(analysis_data)), size = size)
+train_data <- analysis_data[train_index, ]
+test_data <- analysis_data[-train_index, ]
+
+# Fit Bayesian logistic model on the training set
+model_train <- stan_glm(
+  formula = decision ~ application_type + year + planning_district,
+  data = train_data,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5),
+  prior_intercept = normal(location = 0, scale = 2.5),
+  seed = 853
+)
+
+# Predict on the test set
+predicted_probs <- posterior_epred(model_train, newdata = test_data)
+predicted_class <- ifelse(rowMeans(predicted_probs) > 0.5, 1, 0)
+
+# Calculate classification accuracy
+accuracy_train <- mean(predicted_class == train_data$decision)
+accuracy_test <- mean(predicted_class == test_data$decision)
+accuracy_train
+accuracy_test
+
+
+#### Lint and style the code ####
+lint(filename = here("scripts/06-model_validation.R"))
+style_file(path = here("scripts/06-model_validation.R"))

### 2024-12-02 paper/paper.qmd
Upload accuracy result rds file

@@
 
+This study is important because it addresses an area of urban development that has direct implications for policy-making and planning efficiency.
+By analyzing approval patterns, the research helps clarify how procedural factors, geographic contexts, and temporal variations affect outcomes, providing a deeper understanding of the mechanisms at work in planning decisions.
+This understanding is essential for fostering accountability and ensuring that urban growth aligns with community needs and regulatory frameworks.
+The findings also serve as a foundation for improving processes and reducing inconsistencies, contributing to more equitable and transparent urban governance.
+
 The remainder of this paper is structured as follows.
@@
 Then, @sec-result highlights the prediction outcome and results by model summary and plots.
-Finally, @sec-dis includes discussions based on the results, linking it to the real world and also states some strengths and weakness of the study.
+Finally, @sec-dis includes discussions based on the results, linking it to the real world and also states some weakness of the study.
 
@@
 
-Applications to the Committee of Adjustment are typically submitted by property owners, developers, or their authorized agents (e.g., architects or planners).
+Applications to the Committee of Adjustment are typically submitted by property owners, developers, or their authorized agents (e.g.
+architects or planners).
 Eligibility to submit an application requires ownership or legal rights to the property in question.
-For some application types, such as consents, additional prerequisites include demonstrating compliance with existing zoning bylaws or providing documentation showing that the proposed subdivision aligns with city planning objectives.
+There are two types of application, minor variance (MV) and consents (CO).
+Consent applications address land subdivision for developments or sales, while Minor Variance applications request small adjustments to zoning regulations like setbacks or height limits.
+For consent applications, additional prerequisites include demonstrating compliance with existing zoning bylaws or providing documentation showing that the proposed subdivision aligns with city planning objectives.
 Minor variance applications, on the other hand, require applicants to show that the requested adjustments are “minor” in nature, do not conflict with broader zoning intentions, and maintain compatibility with the surrounding area.
@@
 
+Applications to the Committee of Adjustment can be submitted for properties located in one of Toronto’s defined planning districts: Etobicoke York, North York, Scarborough, and Toronto East York.
+Each district represents a distinct geographic area within the city, characterized by varying levels of urban density, development patterns, and zoning requirements.
+The choice of district is determined by the location of the property for which the application is being made.
+
+For the submission year, the dataset captures applications over a range of years, reflecting historical patterns and trends in urban development.
+Applications from earlier years may correspond to a period of slower development or more stringent zoning regulations, while more recent submissions could align with economic growth, population increases, or shifts in city planning priorities.
+For instance, years coinciding with major municipal initiatives or economic booms may see spikes in application numbers.
+By including the year of submission, the dataset provides a temporal dimension that connects development trends to broader economic and policy contexts, such as changes in zoning laws or housing demand.
+
 The data collection process begins with property owners or developers submitting applications through a centralized system.
+Application based surveys are collected.
 Each application includes essential details, which are systematically verified and recorded by administrative staff.
@@
 prior comparisons, and convergence diagnostics, are provided in the appendix for further examination.
-Below, the focus is on other key validation methods:
-
-1. **Out-of-Sample Testing**
+Below, the focus is on key validation method of **out of sampling**:
 
@@
 The model was trained on the training set, and predictions were made on the test set.
-Classification accuracy was calculated as the proportion of correctly classified outcomes in the test set.
+Classification accuracy was calculated as the proportion of correctly classified outcomes in the train set and test set.
 This approach assesses the model's generalizability.
 
-2. **Confusion Metrix**
-The confusion matrix was computed to evaluate the model's classification performance, providing measures such as accuracy, sensitivity, and specificity.
-This validation assesses the model's predictive capabilities and its performance in correctly classifying approval decisions.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-accuracy
+#| tbl-cap: "Classification Accuracy for Model Validation"
+#| warning: false
+
+# Load the saved results
+accuracy_results <- readRDS(file = here::here("models/accuracy_results.rds"))
+
+# Extract the values
+accuracy_train <- accuracy_results$accuracy_train
+accuracy_test <- accuracy_results$accuracy_test
+
+#### Summary Statistics for Accuracy ####
+accuracy_summary <- tibble(
+  `Train Accuracy` = accuracy_train,
+  `Test Accuracy` = accuracy_test
+)
+
+#### Create a kable table ####
+accuracy_summary %>%
+  kable(
+    format = "html"
+  )
+```
+
+@tbl-accuracy presents the train and test accuracy.
+The model's training accuracy of 66.75% and test accuracy of 66.575% are quite close, indicating that the model generalizes reasonably well between the training and test datasets.
+However, the relatively low accuracy on both datasets suggests the presence of underfitting, where the model fails to capture the complexity of the data and adequately explain the variance in the outcomes.
+This could result from the model being too simplistic, with insufficient predictors or inadequate representation of the data's structure.
+Additionally, potential issues like noisy or unbalanced data, insufficient feature engineering, or a lack of higher-order interactions and non-linearities in the model could contribute to this underfitting.
 
@@
 This assumption may oversimplify real-world interactions between variables, such as how specific planning districts may interact with application types or temporal trends.
+The model's relatively low accuracy for both the training and test datasets suggests underfitting, indicating that it fails to capture the full complexity of the data and may not adequately explain the variance in outcomes.
+This limitation could stem from insufficient predictors, unaccounted non-linear relationships, or interactions among variables.
 Furthermore, the model does not fully account for potential multicollinearity among predictors, which could distort the estimated coefficients.
@@
 
-## Idealized Methodology and Survey
-
-### Overview
+# Idealized Methodology and Survey
+
+## Overview
 This section outlines an idealized methodology and survey design to supplement the observational dataset used in this study.
@@
 
-### Sampling Approach
-
-#### Target Population
+## Sampling Approach
+
+### Target Population
 The target population comprises stakeholders involved in the Committee of Adjustment's decision-making process, including:
@@
 
-#### Sampling Frame
+### Sampling Frame
 The sampling frame would be constructed using publicly available application records and contact information from municipal directories.
@@
 
-#### Sample Size
+### Sample Size
 Using stratified random sampling, the sample size would aim for approximately 3000 respondents, ensuring statistical validity while considering feasibility.
 
-#### Stratified Random Sampling
+### Stratified Random Sampling
 Stratified random sampling enhances the accuracy of survey estimates because it ensures proportional representation of different demographic groups.
@@
 
-#### Trade-offs
+### Trade-offs
 While stratified random sampling ensures proportional representation and enhances the precision of survey estimates, it is not without challenges.
@@
 
-### Data Validation
+## Data Validation
 Survey responses would be validated through cross-referencing with the cleaned dataset. For instance:
@@
 
-### Survey
+## Survey
 
@@
 
-#### Survey Structure
+### Survey Structure
 
@@
 
-### Surevey Design Consideration
+## Surevey Design Consideration
 
@@
 
-### Strength
+## Strength
 
@@
 
-### Weakness
+## Weakness
 

### 2024-12-02 scripts/06-model_validation.R

@@ -1,10 +1,11 @@
 #### Preamble ####
-# Purpose: Bayesian Logisitic Regression Model
+# Purpose: Bayesian Logisitic Regression Model Validation
 # Author: Angel Xu
-# Date: 23 November 2024
+# Date: 30 November 2024
 # Contact: anjojoo.xu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: run 01-dowload_data.R, 03-clean_data.R to get the dataset
+# Pre-requisites: run 01-dowload_data.R, 03-clean_data.R, and 05-model_data
+# to get the dataset and the model
 # Any other information needed? None
 
 
@@ -12,6 +13,7 @@
 set.seed(853)
 library(dplyr)
 library(arrow)
+library(rstanarm)
 library(lintr)
 library(styler)
 library(here)
@@ -48,6 +50,10 @@ accuracy_test <- mean(predicted_class == test_data$decision)
 accuracy_train
 accuracy_test
 
+# Save accuracy_train and accuracy_test to an RDS file
+saveRDS(list(accuracy_train = accuracy_train, accuracy_test = accuracy_test),
+  file = "models/accuracy_results.rds"
+)
 
 #### Lint and style the code ####
 lint(filename = here("scripts/06-model_validation.R"))

### 2024-12-03 other/datasheet/datasheet_template.qmd
Upload datasheet

---
title: "Datasheet for Committee of Adjustment Applications Dataset"
subtitle: "Understanding more about the dataset"
author: 
  - Angel Xu
thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

The dataset discussed in this datasheet is 'About Committee of Adjustment Applications' [@committee_application].

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze urban planning applications submitted to Toronto’s Committee of Adjustment. It identifies patterns influencing approvals based on application type, submission year, and planning district, addressing the gap in urban planning research for transparent decision-making.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was compiled and maintained by the City of Toronto's Committee of Adjustment, published through the City of Toronto Open Data Portal.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - It maintained as part of municipal operations by the City of Toronto, likely funded by local government resources.
4. *Any other comments?*
    - None

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset represent formal applications submitted to the City of Toronto’s Committee of Adjustment. Each instance corresponds to a unique application, detailing the specific request for zoning variances or land use consents.
	- Multiple types of instances in the dataset primarily include: 
	  - Applications for Minor Variances (MV): Requests for small adjustments to zoning bylaws, such as building height limits or property setbacks.
	  - Applications for Consents (CO): Requests related to land subdivision or adjustments required for property transactions or development.
	- In addition to these core application types, the dataset also includes related information about the:
	  - Applicants: Individuals or entities submitting the applications, such as property owners, developers, or authorized agents.
	  - Properties: Details about the location and zoning characteristics of the properties associated with the applications.
	  - Decisions: The outcomes of the applications (e.g., approved, refused) once processed by the Committee of Adjustment.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are 3559 instances in total and will continue increase as new data is updated.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*

	- It contains all recorded applications submitted to the City of Toronto’s Committee of Adjustment during the specified timeframe. It is a comprehensive record rather than a sample, systematically maintained by the city to include all relevant applications processed by the committee.
	- While the dataset encompasses all formal applications within its defined scope, it represents a subset of the broader context of land-use adjustments in Toronto. This broader context may include informal requests resolved without a formal application, adjustments handled by other municipal departments, and activities not requiring Committee of Adjustment approval. Additionally, applications outside the specified timeframe are excluded.
	- The dataset is representative of formal applications submitted to the Committee of Adjustment, reflecting geographic and temporal variations across all planning districts in Toronto. However, its focus on formal submissions means it does not account for informal or administrative adjustments. The representativeness of the dataset is validated through its complete coverage of applications within the committee’s jurisdiction, ensuring systematic inclusion of all eligible instances during the specified period.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance in the raw dataset represents a formal application submitted to the City of Toronto’s Committee of Adjustment and consists of structured data in the form of features. These features capture key attributes of each application, including:
	  - Application Details: Type of application (Minor Variance or Consent), submission year, and purpose.
	  - Property Information: Geographic location (planning district), zoning designations, and property identifiers.
	  - Administrative Data: Applicant details (e.g. owner, developer, or agent), decision status (e.g. approved or refused), and times for submission and resolution.
	  - Descriptive Features: Project descriptions and specific requests related to zoning variances or consents.

5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, each instance has a label or target associated with it, the decision outcome of the application.
	- The decision outcome is a variable indicating the final application decision by the Committee of Adjustment. This label reflects the final resolution of the application process, determined after administrative review, public hearings, and committee deliberation. It serves as the primary target for analysis in the cleaned dataset, where it is further processed into a binary format (1 for approved, 0 for refused) to enable predictive modeling and statistical analysis.
  - The decision outcome encapsulates the core purpose of the dataset: to analyze patterns and factors influencing urban planning decisions in Toronto.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some instances in the raw dataset contain missing information due to incomplete submissions or administrative oversights. Missing data may include applicant details, project descriptions, property information, or time spots, often resulting from errors during data entry or non-mandatory fields in the submission system. External factors, such as unregistered properties or unclear zoning bylaws, also contribute to these gaps.
	- These missing data points reflect the challenges of maintaining comprehensive administrative records. The cleaned dataset addressed these issues through preprocessing, such as filtering and reconstructing variables, ensuring the data was complete and reliable for analysis.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Relationships between individual instances are not explicitly represented in the raw dataset. Each instance corresponds to an independent application submitted to the Committee of Adjustment, with no direct links or interactions recorded between applications.
	- However, implicit relationships can be inferred through shared attributes such as planning districts, submission years, or application types. For example, multiple applications from the same district may reflect localized development trends or zoning challenges. These relationships are not explicitly encoded but can be analyzed through groupings or aggregations of shared variables in the dataset.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- There are no predefined splits in the raw dataset while a data split is applied on the cleaned dataset for model validation. The dataset is divided into training (80%) and testing (20%) subsets.
	- The rationale for this split is to use the training set for model building and parameter estimation, while the testing set evaluates the model’s performance on unseen data. This ensures the model generalizes well and avoids overfitting.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Self-contained.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, any personally identifiable information about applicants, such as names, contact details, or specific property addresses, is either excluded or anonymized in the dataset.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No. The dataset does not explicitly identify sub-populations based on demographic attributes such as age or gender. Instead, sub-populations are represented through administrative and geographic groupings relevant to urban planning decisions.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.

16. *Any other comments?*
	- None.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data for each instance was primarily directly observable and recorded as part of the formal application process to the City of Toronto’s Committee of Adjustment. Key data points, such as application type, geographic location, and submission dates, were submitted directly by applicants through a centralized online system or physical forms. Decision outcomes were added by committee staff following administrative review and public hearings.
	- Some data, such as zoning compliance or project descriptions, can be considered reported by subjects, as applicants provide this information when submitting their applications. These submissions are validated by committee staff to ensure they align with zoning regulations and meet the submission requirements.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected through a centralized online application system maintained by the City of Toronto’s Committee of Adjustment. Applicants submit their applications via this digital platform or through physical forms, which are then digitized by administrative staff.
	- Administrative staff manually input application details from physical submissions and verify the accuracy of online submissions. Besides, submitted data is stored in a centralized digital database that standardizes and organizes records, ensuring consistency across entries. Submission and resolution dates are automatically records in the system.
	- For validation, submitted applications are reviewed by committee staff to ensure compliance with zoning bylaws and completeness of information. Missing or unclear details are flagged for clarification with applicants. Additionally, the centralized system enforces mandatory fields during the submission process, reducing the likelihood of incomplete records. The database is also maintained and updated regularly to reflect accurate and current information, ensuring its reliability for public use and analysis.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is not a sample。

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data collection process was conducted by administrative staff and officials at the City of Toronto’s Committee of Adjustment. These staff members are municipal employees who are compensated as part of their regular salaries funded by the city government. They were responsible for verifying, recording, and maintaining the data submitted by applicants.
	- Applicants submitted their applications voluntarily as part of the formal urban planning process. They were not compensated for providing their data, as the submission is a required step for obtaining zoning variances or consents. Application fees paid by the applicants help fund the administrative operations of the Committee of Adjustment.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The raw dataset includes applications submitted from 1999 to the present. This timeframe matches the creation timeframe of the data, as the dataset records information at the time applications are submitted, reviewed, and decided upon. The times for submissions, hearings, and decisions accurately reflect the actual events as they occurred, ensuring no significant temporal discrepancies between the creation and collection of data.
	- The cleaned dataset filtered out earlier years, which have sparse data, to improve analytical reliability, but the timeframe still aligns with the period during which the applications were originally created.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No. The data was collected and maintained by the City of Toronto’s Committee of Adjustment as part of routine administrative operations. Since the dataset primarily contains anonymized and aggregated administrative data without sensitive personal identifiers, it does not fall under categories that typically require ethical review.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was collected directly from individuals and entities submitting applications to the City of Toronto’s Committee of Adjustment.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Yes, individuals submitting applications to the City of Toronto’s Committee of Adjustment were notified about the data collection. The submission process, whether online or in-person, includes a notification that the information provided will be used for processing the application and may be recorded in public records as part of municipal governance procedures.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Individuals submitting applications to the City of Toronto’s Committee of Adjustment implicitly consented to the collection and use of their data as part of the formal application process. Consent is typically provided by agreeing to terms of service or privacy statements for online submissions or signing disclaimers included on physical forms.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- There is no indication that individuals submitting applications to the City of Toronto’s Committee of Adjustment are provided with a formal mechanism to revoke their consent after submission.

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.

12. *Any other comments?*
	- None.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- No, not mentioned.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, can be accessed in `data/01-raw_data`.

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- R [@citeR] was used. For cleaning process of used analysis dataset in this paper, it can be accessed in `scripts/03-clean_data.R`.

4. *Any other comments?*
	- None.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, the dataset has been used to analyze factors influencing urban planning decisions in Toronto by Angel Xu. No other use of the dataset is found.

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.

3. *What (other) tasks could the dataset be used for?*
	- The dataset has the potential to support various tasks related to urban planning, governance, and policy analysis. It could be used for predictive modeling to estimate application outcomes based on geographic, temporal, and contextual factors, providing insights into decision-making trends. Additionally, the dataset could be applied to process optimization by identifying inefficiencies in the application review and approval process and suggesting administrative improvements. It could also be used to explore spatial analysis to uncover geographic disparities in approval rates, highlighting districts with higher or lower likelihoods of approval and identifying potential inequities or regional planning challenges.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Yes, the dataset reflects historical decision-making processes, which may include systemic biases. Using this dataset without accounting for such factors could perpetuate unfair treatment or stereotyping of specific regions or application types.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- Yes, the dataset should not be used to identify or profile individual applicants since it does not contain personal identifiers and was not designed for such purposes. The dataset should not be used to draw conclusions about zoning or planning decisions in regions outside Toronto as well.

6. *Any other comments?*
	- None.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, the dataset is distributed publicly through the City of Open Data Toronto Portal [@opendatatoronto].

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed through the City of Open Data Toronto Portal [@opendatatoronto] as downloadable files (can be download as CSV, JSON or XML format). Users can access the data directly from the portal’s website, where it is periodically updated and maintained by the city. However, the dataset does not have a DOI.

3. *When will the dataset be distributed?*
	- It is updated daily to reflect new applications and decisions, ensuring that the data remains current and relevant. Users can access the dataset at any time.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Yes, the dataset is distributed under the City of Toronto’s Open Data License, which allows public use with certain restrictions. The license permits users to copy, modify, publish, and share the data, provided that proper attribution is given to the City of Toronto. It prohibits using the data in ways that misrepresent its source or for unlawful purposes.
	- No fees is associated with accessing or using the dataset under this license.

5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.

7. *Any other comments?*
	- None.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset is supported, hosted, and maintained by the City of Toronto, specifically through its Open Data Portal [@opendatatoronto]. The City’s municipal departments are responsible for collecting, verifying, updating, and publishing the data.
	- Regular updates are performed to ensure accuracy and reflect the latest submissions and decisions by the Committee of Adjustment.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The owner, curator, or manager of the dataset can be contacted through the City of Toronto’s Open Data team or the Committee of Adjustment.
	- For general inquiries about the dataset, users can email the City of Toronto's Open Data team at [opendata@toronto.ca](opendata@toronto.ca). For specific questions about the Committee of Adjustment applications, additional contact details are available on the [Committee of Adjustment website](https://www.toronto.ca/city-government/planning-development/committee-of-adjustment/).

3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, the dataset is refreshed dailty while updated when new applications occur. Updates are managed by the City of Toronto’s Open Data team in collaboration with the Committee of Adjustment. Updates are communicated to dataset consumers through the Open Data Toronto Portal [@opendatatoronto].

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset relates to formal applications rather than personal data. Thus, no explicit limits on the retention of data are specified.

6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- There is no formal mechanism for external parties to directly extend, augment, or contribute to the dataset. However, others can contact the Open Data team at [opendata@toronto.ca](opendata@toronto.ca) if they wish to provide feedback or suggest improvements.

8. *Any other comments?*
	- None.

\newpage

# References

### 2024-12-03 paper/paper.qmd
Upload datasheet

@@
 subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
-author: Angel Xu
+author: Victoria Zheng
 thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
@@
   geom_bar(fill = "lightblue", color = "black") +
-  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
   labs(
@@
   )
-
 ```
@@
   summarize(
-    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add '%' symbol
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
   ) %>%
@@
   geom_bar(fill = "lightgreen", color = "black") +
-  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
   labs(
@@
   summarize(
-    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%") # Add %
+    Percentage = paste0(round((n() / nrow(analysis_data)) * 100, 1), "%")
   ) %>%
-  mutate(Application_Type = ifelse(application_type == "MV", "Minor Variance (MV)", "Consent (CO)")) %>%
+  mutate(Application_Type = ifelse(application_type == "MV",
+    "Minor Variance (MV)", "Consent (CO)"
+  )) %>%
   select(Application_Type, Percentage)
@@
   geom_bar(fill = "skyblue", color = "black") +
-  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
   labs(
@@
   geom_bar(fill = "coral", color = "black") +
-  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
   labs(
@@
     axis.text = element_text(size = 12),
-    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate district labels for readability
-  )
-
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
 ```
@@
 These predictors are naturally suited to a logistic framework, and the Bayesian approach accommodates any inherent variability in their effects.
+
+`Mostly Harmless Econometrics` (@angrist2009mostly) addresses how to handle variability in time-series and cross-sectional data, focusing on ensuring robust and interpretable results.
+The use of submission year and planning district as key predictors aligns with Angrist and Pischke's guidance on analyzing temporal and geographic variation.
+Their discussion on the importance of controlling for time trends and location-specific effects validates the inclusion of these variables in the Bayesian logistic regression model.
 
@@
   theme_minimal()
-
 ```
@@
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-subcap: ["Posterior prediction check",
+#| "Comparing the posterior with the prior"]
 
@@
 
-This method reduces sampling variability, enhances representation, and provides more accurate estimates compared to simple random sampling, as the unique characteristics of each subgroup are reflected in the final sample.
+`Sampling: Design and Analysis` (@lohr2021sampling) explores how stratified sampling reduces variability by dividing populations into homogeneous groups, enabling more precise estimates and better resource allocation.
+By applying these principles, the paper ensures that the analysis accounts for heterogeneity in geographic and temporal patterns of planning applications, enabling robust conclusions about decision-making trends.
+The methodology follows Lohr’s recommendation to tailor strata definitions to the study's objectives, enhancing the reliability of the results.
+
+Besides, `Sampling` (@thompson2012sampling) provides a detailed guide to modern sampling techniques, emphasizing the importance of tailoring approaches to study objectives.
+His discussion on stratified sampling highlights its benefits in improving precision and reducing bias by dividing populations into homogeneous subgroups.
+This aligns with the choice of using planning districts, application types, and submission years as strata to ensure balanced representation and capture temporal and spatial variability in urban planning decisions.
+Thompson’s principles validate the proportional sampling approach adopted here, supporting its effectiveness in generating reliable analyses.
 
@@
 - The survey is estimated to take about 5 minutes to complete.
-- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
+- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).
 
@@
 
-If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
+If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).
 

### 2024-12-03 paper/paper.qmd
Upload datasheet name

@@
 subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
-author: Victoria Zheng
+author: Angel Xu
 thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
@@
 - The survey is estimated to take about 5 minutes to complete.
-- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).
+- If you have any questions or concerns, feel free to reach out to me at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
 
@@
 
-If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Victoria Zheng).
+If you have any questions or would like to learn more about this study, please don’t hesitate to contact our research team at [anjojoo.xu@mail.utoronto.ca](anjojoo.xu@mail.utoronto.ca) (Angel Xu).
 

### 2024-12-03 paper/paper.qmd
Revise paper

@@
 Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews.
-These findings enhance understanding of urban planning decision-making processes, offering insights into temporal trends, geographic variability, and areas for equitable policy adjustments."
+These findings enhance understanding of urban planning decision-making processes, offering perceptions into temporal trends, geographic variability, and areas for equitable policy adjustments."
 format: pdf
@@
 By analyzing variables such as the type of application (minor variance or consent), the year of submission, and the planning district, this study seeks to understand how these elements shape the decision-making process.
-The primary objective is to determine the causal effects of these factors on approval outcomes, thereby offering insights into the dynamics of urban planning and regulatory processes.
+The primary objective is to determine the causal effects of these factors on approval outcomes, thereby offering perceptions into the dynamics of urban planning and regulatory processes.
 
@@
 planning districts), and the timing of submission.
-Understanding these decisions provides insight into urban planning dynamics and the regulatory challenges faced by a growing city like Toronto.
+Understanding these decisions provides awareness of urban planning dynamics and the regulatory challenges faced by a growing city like Toronto.
 
@@
 By distilling real-world phenomena—such as zoning adjustments and land use approvals—into discrete variables, the dataset enables systematic analysis of trends and relationships within Toronto’s regulatory framework.
-This structured approach ensures that the dataset is both practical for analysis and representative of the nuanced processes underlying urban development.
+This structured approach ensures that the dataset is both practical for analysis and representative of the detailed processes underlying urban development.
 
@@
   geom_bar(fill = "lightblue", color = "black") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -1.0) +
   labs(
@@
   ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
   theme_minimal() +
@@
   geom_bar(fill = "lightgreen", color = "black") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -1.0) +
   labs(
@@
 The year variable represents the year of submission for each application in the dataset.
-It provides insight into temporal trends in applications submitted to the Committee of Adjustment.
+It provides awareness into temporal trends in applications submitted to the Committee of Adjustment.
 This variable is instrumental in identifying changes in the volume of applications over time, which may reflect policy shifts, urban development trends, or external factors like economic conditions.
@@
   geom_bar(fill = "skyblue", color = "black") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.8) + # Move labels higher
   labs(
@@
   ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + # Add extra space above bars
   theme_minimal() +
@@
 As approval rates may have varied over time due to changing city planning policies or committee priorities.
-Besides, temporal trends may reveal shifts in the proportion of Minor Variances (MV) versus Consents (CO), reflecting evolving development demands.
+Besides, temporal trends may highlight shifts in the proportion of Minor Variances (MV) versus Consents (CO), reflecting evolving development demands.
 
@@
 The city is divided into four planning districts, each with its own development characteristics and zoning regulations.
-These districts are crucial for understanding spatial trends in application submissions and outcomes, as local policies and priorities can influence decision-making processes.
+These districts are pivotal for understanding spatial trends in application submissions and outcomes, as local policies and priorities can influence decision-making processes.
 
@@
   geom_bar(fill = "coral", color = "black") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -1.0) + # Move labels higher
   labs(
@@
   ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) + # Add extra space above bars
   theme_minimal() +
@@
     axis.text = element_text(size = 12),
-    axis.text.x = element_text(angle = 45, hjust = 1)
+    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis text
   )
@@
 The resulting posterior distributions provide updated beliefs about the coefficients, incorporating both prior knowledge and evidence from the data.
-This approach allows for robust inference, offering a clear quantification of parameter uncertainty and supporting more nuanced interpretations of the predictors' effects.
+This approach allows for robust inference, offering a clear quantification of parameter uncertainty and supporting more detailed interpretations of the predictors' effects.
 
@@
 This choice is driven by the binary nature of the response variable, the model's ability to incorporate prior knowledge, and its probabilistic framework, which quantifies uncertainty more effectively than frequentist alternatives.
-Logistic regression naturally constrains predicted probabilities to the [0,1] interval, ensuring interpretability as likelihoods of approval, while the Bayesian approach enhances flexibility and provides richer insights through posterior distributions.
+Logistic regression naturally constrains predicted probabilities to the [0,1] interval, ensuring interpretability as likelihoods of approval, while the Bayesian approach enhances flexibility and provides richer revelations through posterior distributions.
 
@@
 This Bayesian model relies on several assumptions that ensure the validity of its predictions and the transparency of its coefficients.
-While these assumptions are generally less restrictive than those for simpler models like linear regression, they remain critical to the robustness of the results.
+While these assumptions are generally less restrictive than those for simpler models like linear regression, they remain important to the robustness of the results.
 The key assumptions are listed below:
@@
 The result highlights how the predictors influence the likelihood of application approval while controlling for other variables.
-The coefficients, presented as log-odds, provide insights into both the direction and magnitude of each predictor’s impact on the approval decision.
+The coefficients, presented as log-odds, provide observations into both the direction and magnitude of each predictor’s impact on the approval decision.
 
@@
 
-The year variables reveal significant temporal variations in approval likelihood, using 2016 as the reference year.
+The year variables display significant temporal variations in approval likelihood, using 2016 as the reference year.
 These coefficients highlight how approval dynamics have shifted over time.
@@
 
-In summary, the model reveals key patterns in how application type and submission year influence the likelihood of approval.
+In summary, the model presents key patterns in how application type and submission year influence the likelihood of approval.
 Minor variance applications face slightly lower odds of approval compared to consents, while temporal trends indicate an overall increase in approval rates from 2016 to 2020.
-These findings provide a foundation for understanding the dynamics of approval decisions and highlight areas for further exploration, such as geographic variability or interactions between predictors.
+These findings provide a foundation for understanding the dynamics of approval decisions and highlight areas for further investigation, such as geographic variability or interactions between predictors.
 
@@
 This paper investigates the factors influencing application approvals submitted to the City of Toronto's Committee of Adjustment, using a Bayesian logistic regression model.
-By analyzing predictors such as application type, submission year, and planning district, the study provides insights into the dynamics of urban planning decisions.
+By analyzing predictors such as application type, submission year, and planning district, the study provides findings into the dynamics of urban planning decisions.
 The model captures temporal trends, spatial variability, and procedural differences, offering a structured understanding of approval outcomes.
@@
 One key finding is the temporal trend in approval rates.
-The analysis reveals that approval odds have increased steadily over time, peaking in 2018.
+The analysis exhibits that approval odds have increased steadily over time, peaking in 2018.
 This trend may reflect evolving planning policies, greater efficiency in committee processes, or a shift in zoning priorities.
@@
 Conversely, Scarborough has the lowest odds of approval, suggesting stricter zoning enforcement or less alignment with development objectives in this area.
-This geographic disparity underscores the influence of local contexts and priorities on urban planning decisions, revealing how different districts balance growth and zoning compliance.
+This geographic disparity underscores the influence of local contexts and priorities on urban planning decisions, manifesting how different districts balance growth and zoning compliance.
 
@@
 
-While this paper provides insights into the factors influencing application approvals, there exists several limitations.
+While this paper provides findings into the factors influencing application approvals, there exists several limitations.
 These weaknesses cover data-related issues, methodological constraints, and broader contextual considerations.
@@
 1. **Data Limitations**
-The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain critical variables.
+The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain important variables.
 Key contextual factors, such as community objections, project complexity, or the presence of political or economic pressures, are not captured.
@@
 For example, public hearings, lobbying efforts, and input from local communities likely play a significant role in shaping decisions but are absent from the dataset.
-This simplification means that the model captures only part of the story, potentially overlooking nuanced dynamics in urban planning and governance.
+This simplification means that the model captures only part of the story, potentially overlooking sophisticated dynamics in urban planning and governance.
 
@@
 
-One critical next step is to expand the dataset by incorporating additional variables that capture contextual and qualitative factors influencing decisions.
-Data on community feedback and objections during public hearings is a good choice since it could provide insights into how local resistance or support impacts approval likelihood.
-Information on project complexity, such as the scale, type, and potential impact of proposals, could reveal patterns in decision-making.
-Besides, access to committee meeting notes or further exploration in meeting minutes would help clarify the reasoning behind specific approvals or refusals.
+One important next step is to expand the dataset by incorporating additional variables that capture contextual and qualitative factors influencing decisions.
+Data on community feedback and objections during public hearings is a good choice since it could show how local resistance or support impacts approval likelihood.
+Information on project complexity, such as the scale, type, and potential impact of proposals, could indicate patterns in decision-making.
+Besides, access to committee meeting notes or further investigation in meeting minutes would help clarify the reasoning behind specific approvals or refusals.
 Furthermore, integrating economic and political context, such as local economic indicators or changes in political leadership, would offer a broader understanding of approval trends.
@@
 Future research could explore nonlinear relationships, such as whether certain predictors have diminishing or exponential effects on approval rates.
-Interaction effects between predictors, such as the interplay between planning districts and application types, could also provide more nuanced insights.
+Interaction effects between predictors, such as the interplay between planning districts and application types, could also provide more subtle observation.
 For instance, minor variances may be treated differently in densely populated urban areas compared to suburban districts.
@@
 
-Despite to future improvements for addressing limitations, further exploration can be continued based on the current result.
+Despite to future improvements for addressing limitations, further examination can be continued based on the current result.
 The temporal trends identified in this paper suggest shifts in approval rates over time, but the reasons behind these changes remain unexplored.
@@
 Spatial regression models could account for spatial autocorrelation and clustering effects in planning districts, while network analysis could study relationships between districts, such as spillover effects of development activity.
-Additionally, analyzing proximity to key infrastructure, such as transit hubs or commercial centers, could provide insights into spatial patterns of decision-making.
+Additionally, analyzing proximity to key infrastructure, such as transit hubs or commercial centers, could provide perceptions into spatial patterns of decision-making.
 
@@
 In summary, future research should prioritize richer data integration, explore more sophisticated modeling techniques, and broaden the scope of analysis to other regions and temporal contexts.
-These steps would not only enhance academic understanding of urban planning decisions but also provide actionable insights for policymakers, developers, and urban planners working toward effective and equitable processes.
+These steps would not only enhance academic understanding of urban planning decisions but also provide actionable perceptions for policymakers, developers, and urban planners working toward effective and equitable processes.
 
@@
 This comparison highlights how the data influenced parameter estimates.
-Large shifts from the prior to the posterior, such as seen for certain planning districts and years, suggest strong data-driven updates.
+Large shifts from the prior to the posterior, such as seen for certain planning districts and years, suggest strong empirical updates.
 Conversely, parameters where the posterior closely resembles the prior indicate minimal influence from the observed data.
@@
 
-@fig-stanareyouokay-1 is a trace plot.
-It shows he sampled values for each chain of the Markov Chain Monte Carlo (MCMC) simulation over iterations.
+@fig-trace presents trace plots.
+It shows the sampled values for each chain of the Markov Chain Monte Carlo (MCMC) simulation over iterations.
 The chains appear to mix well and explore the parameter space without any evident trends or stickiness, suggesting convergence.
@@
 
-@fig-stanareyouokay-2 is a Rhat plot.
+@fig-rhat is a Rhat plot.
 It evaluates the Gelman-Rubin convergence diagnostic.
@@
 #| warning: false
-#| label: fig-stanareyouokay
+#| label: fig-trace
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Application type MV", "Trace plot of year 2017", "Trace plot of year 2018", "Trace plot of year 2019", "Trace plot of year 2020", "Trace plot of year 2021", "Trace plot of year 2022", "Trace plot of year 2023", "Trace plot of year 2024", "Trace plot of planning district NorthYork", "Trace plot of planning district Scarborough", "Trace plot of planning district EastYork"]
+#| layout-ncol: 3
+
+plot(model, "trace", "(Intercept)")
+plot(model, "trace", "application_typeMV")
+plot(model, "trace", "year2017")
+plot(model, "trace", "year2018")
+plot(model, "trace", "year2019")
+plot(model, "trace", "year2020")
+plot(model, "trace", "year2021")
+plot(model, "trace", "year2022")
+plot(model, "trace", "year2023")
+plot(model, "trace", "year2024")
+plot(model, "trace", "planning_districtNorth York")
+plot(model, "trace", "planning_districtScarborough")
+plot(model, "trace", "planning_districtToronto East York")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-rhat
+#| fig-cap: "Rhat plot"
 #| layout-ncol: 1
-
-plot(model, "trace")
 
@@
 Afterward, random samples are drawn independently from each subgroup, ensuring proportional representation.
-In cases where smaller groups are critical to the study, adjustments can be made to ensure adequate representation.
+In cases where smaller groups are important to the study, adjustments can be made to ensure adequate representation.
 Finally, the selected samples from each subgroup are combined to create the overall sample, ensuring that all significant categories are included.
@@
 I appreciate your participation in this survey, which seeks to better understand the factors influencing urban planning decisions and application outcomes in Toronto.
-Your responses are crucial for enhancing the research and contributing to insights that could inform future urban planning processes.
+Your responses are pivotal for enhancing the research and contributing to perceptions that could inform future urban planning processes.
 
@@
 
-Thank you for your valuable contribution! As a token of our appreciation, each participant will receive $5 upon completion of the survey.
+Thank you for your significant contribution! As a token of our appreciation, each participant will receive $5 upon completion of the survey.
 
@@
 
-Do you have any further comments or insights about the factors that might affect the application decision?
+Do you have any further comments or awareness about the factors that might affect the application decision?
 
@@
 
-Do you have any further comments or insights about the application process?
+Do you have any further comments or awareness about the application process?
 
@@
 Your participation is essential in helping us gain a deeper understanding of the factors influencing urban planning decisions in Toronto.
-Your insights will contribute to research aimed at improving zoning application processes and understanding decision dynamics.
+Your awareness will contribute to research aimed at improving zoning application processes and understanding decision dynamics.
 
@@
 
-As a token of our gratitude, each participant will receive a $5 reward shortly after completing the survey. Thank you again for your valuable contribution!
+As a token of our gratitude, each participant will receive a $5 reward shortly after completing the survey. Thank you again for your significant contribution!
 
@@
 Efforts were also made to limit the survey's length to encourage participation while ensuring sufficient data for meaningful analysis.
-A balance was struck between closed-ended questions, which facilitate quantitative analysis, and open-ended questions, which capture nuanced qualitative insights.
+A balance was struck between closed-ended questions, which facilitate quantitative analysis, and open-ended questions, which capture intricate qualitative findings.
 Additionally, the survey design accounted for ethical considerations, such as maintaining confidentiality and emphasizing voluntary participation.
@@
 Stratified random sampling ensures that the results are representative of the population, reducing sampling bias.
-By aligning strata with key demographic and process variables, the survey facilitates detailed subgroup analysis, enhancing the depth of insights.
+By aligning strata with key demographic and process variables, the survey facilitates detailed subgroup analysis, enhancing the depth of discovery.
 The inclusion of both quantitative and qualitative questions allows for an overall understanding of the factors influencing urban planning decisions.
@@
 The survey also addresses gaps in the cleaned dataset, such as perceptions of community feedback, challenges during the application process, and external influences on decisions.
-These insights complement the quantitative data, offering a more holistic view of the decision-making process.
+They complement the quantitative data, offering a more holistic view of the decision-making process.
 
@@
 Furthermore, the survey results are context-specific, reflecting Toronto’s unique regulatory framework and urban planning dynamics, which may limit the generalizability of findings to other cities or contexts.
-Despite these challenges, the survey is a valuable tool for supplementing the dataset and addressing research objectives.
+Despite these challenges, the survey is a useful tool for supplementing the dataset and addressing research objectives.
 

### 2024-12-04 paper/paper.qmd
Paper Revision

@@
 abstract: "This paper analyzes urban planning application approvals submitted to Toronto's Committee of Adjustment from 2016 to 2024, focusing on key predictors such as application type, submission year, and planning district using Bayesian logistic regression.
-The analysis shows that consents are more likely to be approved than minor variances, with approval rates generally increasing over time, particularly from 2017 onward, though fluctuations are observed in certain years.
+The analysis shows that Consents are more likely to be approved than Minor Variances, with approval rates generally increasing over time, particularly from 2017 onward, despite fluctuations in certain years.
 Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews.
-These findings enhance understanding of urban planning decision-making processes, offering perceptions into temporal trends, geographic variability, and areas for equitable policy adjustments."
+These findings enhance understanding of urban planning decision-making processes, offering implications into temporal trends, geographic variability, and areas for equitable policy adjustments."
 format: pdf
@@
 library(knitr)
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+library(arrow)
+
+analysis_data <- read_parquet(here::here(
+  "data/02-analysis_data/analysis_data.parquet"))
 ```
@@
 This study is important because it addresses an area of urban development that has direct implications for policy-making and planning efficiency.
-By analyzing approval patterns, the research helps clarify how procedural factors, geographic contexts, and temporal variations affect outcomes, providing a deeper understanding of the mechanisms at work in planning decisions.
+By analyzing approval patterns, the research clarifies how procedural factors, geographic contexts, and temporal variations affect outcomes, providing a deeper understanding of the mechanisms at work in planning decisions.
 This understanding is essential for fostering accountability and ensuring that urban growth aligns with community needs and regulatory frameworks.
@@
 
-This study aims to estimate the factors influencing the approval decisions for urban planning applications in Toronto.
+This study aims to identify the factors influencing approval decisions for urban planning applications in Toronto.
 By analyzing variables such as the type of application (minor variance or consent), the year of submission, and the planning district, this study seeks to understand how these elements shape the decision-making process.
-The primary objective is to determine the causal effects of these factors on approval outcomes, thereby offering perceptions into the dynamics of urban planning and regulatory processes.
+The primary objective is to examine the causal effects of these factors on approval outcomes, thereby offering analysis of the dynamics of urban planning and regulatory processes.
 
@@
 
-Our data [@committee_application] is extracted from opendatatoronto [@opendatatoronto].
+Our data [@committee_application] was extracted from opendatatoronto [@opendatatoronto].
 This dataset focuses on applications submitted to the City of Toronto's Committee of Adjustment, a governing body responsible for reviewing minor variance and consent applications related to land use and development.
@@
 It ensures that deviations from zoning bylaws align with the city’s broader development objectives, maintaining a balance between flexibility for property owners and adherence to long-term planning goals.
-The decisions made by the committee are influenced by various factors, including the type of application, geographic considerations (e.g.
+The decisions made by the Committee are influenced by various factors, including the type of application, geographic considerations (e.g.
 planning districts), and the timing of submission.
-Understanding these decisions provides awareness of urban planning dynamics and the regulatory challenges faced by a growing city like Toronto.
-
-The analysis presented in this paper were conducted using R programming language [@citeR].
-The `tidyverse` packages [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow] were used in the process of data simulation, testing beforehand.
-After the original raw data was downloaded by using `tidyverse` package [@citetidyverse], and `dplyr` package [@citedplyr], data cleaning process was done by using `tidyverse` package [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow].
-The code style is corrected by `here` package [@citehere], `lintr` package [@citelintr], and `styler` package [@citestyler].
-Then, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
-Graphs were made with `ggplot2` package [@citeggplot] and `tidybayes` package [@citetidybayes].
-Tables were constructed with `knitr` package [@citeknitr].
-The model results are then presented by `modelsummary` [@citemodelsummary] package.
-
-Following @tellingstories, this paper uses Bayesian Logistic Model to explore factors that affect the application results.
+Understanding these decisions enhances comprehension of urban planning dynamics and the regulatory challenges faced by a growing city like Toronto.
+
+The analysis presented in this paper was conducted using R programming language [@citeR].
+The `tidyverse` packages [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow] were utilized for data simulation and testing.
+After the original raw data was downloaded using the `tidyverse` package [@citetidyverse], and `dplyr` package [@citedplyr], data cleaning process was carried out using `tidyverse` package [@citetidyverse], `dplyr` package [@citedplyr], and `arrow` package [@citearrow].
+The code style was corrected by `here` package [@citehere], `lintr` package [@citelintr], and `styler` package [@citestyler].
+Subsequently, models were constructed using `tidyverse` package [@citetidyverse], and `rstanarm` [@citerstanarm] package.
+Graphs were generated with `ggplot2` package [@citeggplot] and `tidybayes` package [@citetidybayes].
+Tables were created with `knitr` package [@citeknitr].
+The model results were presented by `modelsummary` [@citemodelsummary] package.
+
+Following @tellingstories, this paper employs a Bayesian Logistic Model to explore factors that affect the application results.
 
 Although several similar datasets with overlapping themes were considered but ultimately not used due to limitations in scope, structure, or relevance.
-For instance, the Toronto Application Information Centre Dataset provides information on broader planning applications like rezoning requests and site plan approvals, but its complexity and inclusion of unrelated application types make it unsuitable for isolating patterns related to minor variances and consents.
+For instance, the Toronto Application Information Centre Dataset provides information on broader planning applications like rezoning requests and site plan approvals, but its complexity and inclusion of unrelated application types made it unsuitable for isolating patterns related to minor variances and consents.
 The Ontario Municipal Board (OMB) Decisions Dataset includes appeal outcomes, but its focus on contested cases introduces bias and omits the routine applications central to this study.
 
-Consequently, focusing on the Committee of Adjustment Applications dataset [@committee_application] stands out for its granularity and relevance, enabling a focused analysis of approval trends and their influencing factors, which directly aligns with the study's goals.
+Consequently, focusing on the Committee of Adjustment Applications dataset [@committee_application] stands out for its granularity and relevance, enabling a focused analysis of approval trends and their influencing factors, which directly align with the study's goals.
 
@@
 The dataset captures information on applications submitted to the City of Toronto’s Committee of Adjustment, providing a structured record of decisions made on zoning variances and consents.
-The dataset’s structure results from Toronto’s well-defined application process.
+This structure reflects Toronto’s well-defined application process.
 When property owners or developers submit applications to the Committee of Adjustment, they provide detailed information about their project, including its location, purpose, and the specific variances or consents requested.
@@
 architects or planners).
-Eligibility to submit an application requires ownership or legal rights to the property in question.
+Eligibility requires ownership or legal rights to the property in question.
 There are two types of application, minor variance (MV) and consents (CO).
@@
 
-Applications to the Committee of Adjustment can be submitted for properties located in one of Toronto’s defined planning districts: Etobicoke York, North York, Scarborough, and Toronto East York.
+Applications to the Committee of Adjustment can be submitted for properties in Toronto’s defined planning districts: Etobicoke York, North York, Scarborough, and Toronto East York.
 Each district represents a distinct geographic area within the city, characterized by varying levels of urban density, development patterns, and zoning requirements.
@@
 
-For the submission year, the dataset captures applications over a range of years, reflecting historical patterns and trends in urban development.
+The dataset captures applications spanning multiple years, reflecting historical patterns and trends in urban development.
 Applications from earlier years may correspond to a period of slower development or more stringent zoning regulations, while more recent submissions could align with economic growth, population increases, or shifts in city planning priorities.
@@
 
-The data collection process begins with property owners or developers submitting applications through a centralized system.
+The data collection process begins when property owners or developers submitting applications through a centralized system.
 Application based surveys are collected.
@@
 Following public hearings, committee decisions (approval or refusal) are added to the database, completing the data entry process.
-The dataset is then maintained and periodically published by the City of Toronto, ensuring it remains accessible and accurate for public and analytical use.
-
-The dataset thus provides a structured lens through which to examine urban planning decisions.
+The dataset is then maintained and regularly published by the City of Toronto, ensuring it remains accessible and accurate for public and analytical use.
+
+The dataset serves as an effective tool for examining urban planning decisions.
 By distilling real-world phenomena—such as zoning adjustments and land use approvals—into discrete variables, the dataset enables systematic analysis of trends and relationships within Toronto’s regulatory framework.
@@
 
-To ensure the dataset was suitable for analysis, several high-level cleaning steps were performed.
-First, unused decision outcomes were filtered out, retaining only applications that were explicitly approved or refused.
+To ensure the dataset was suitable for analysis, several cleaning steps were implemented.
+First, unused decision outcomes were removed, retaining only applications that were explicitly approved or refused.
 The `decision` variable was then converted into a binary format, with `1` representing approvals and `0` representing refusals, simplifying its use in modeling.
@@
 A new variable, `year`, was constructed by extracting the year from the `in_date` field.
-This allowed for a clearer examination of temporal trends while maintaining the granularity of the original data.
-Applications from very early years with minimal counts were excluded to avoid bias and ensure a representative analysis of contemporary decision-making trends.
-
-Additionally, key variables are selected out in cleaned dataset for convenience. More details will be explored in appendix.
+This adjustment enabled a clearer examination of temporal trends while preserving the granularity of the original data.
+Applications from earlier years with minimal counts were excluded to avoid bias and ensure a representative analysis of contemporary decision-making trends.
+
+Additionally, key variables were selected in the cleaned dataset for convenience. More details are provided in the appendix.
 
@@
 
-- Decision: Indicates whether an application was approved or refused.
-- Application Type: Categorizes the application as either a minor variance or a consent.
-- Year: The year when the application was submitted.
-- Planning District: The geographic area within Toronto where the application originated.
+- **Decision**: Indicates whether an application was approved or refused.
+- **Application Type**: Categorizes the application as either a minor variance or a consent.
+- **Year**: The year when the application was submitted.
+- **Planning District**: The geographic area within Toronto where the application originated.
 
@@
 Out of the total applications, 1,207 (66.5%) were approved, while 607 (33.5%) were refused.
-This demonstrates that the majority of applications submitted to the Committee of Adjustment are approved, suggesting that most requests align with zoning regulations or are deemed acceptable following the committee's review process.
+This indicates that most applications are approved, suggesting that most requests align with zoning regulations or are deemed acceptable following the committee's review process.
 
@@
 
-@tbl-decision provides a clear picture of the approval and refusal rates within the dataset, highlighting a relatively high likelihood of approval for submitted applications.
+@tbl-decision summarizes the approval and refusal rates in the dataset, emphasizing a relatively high likelihood of approval for submitted applications.
 
@@
 
-Consent (CO): Requests related to the subdivision of land into separate parcels, often for new developments or property sales.
-
-Minor Variance (MV): Applications seeking small adjustments to zoning regulations, such as building setbacks, height limits, or lot coverage.
-
-These categories represent distinct regulatory processes, with Consents generally requiring more extensive evaluation compared to Minor Variances due to their implications on land use.
+- Consent (CO): Requests related to the subdivision of land into separate parcels, often for new developments or property sales.
+
+- Minor Variance (MV): Applications seeking small adjustments to zoning regulations, such as building setbacks, height limits, or lot coverage.
+
+These categories represent distinct regulatory processes. Consents generally require more extensive evaluation than Minor Variances due to their implications on land use.
 
@@
 Minor Variances account for the majority of applications which reach to 1371 (75.6%), reflecting the routine nature of these requests in urban development.
-In contrast, Consents are less frequent which only have 443 counts (24.4%), likely due to their more specific requirements and potential complexity.
+In contrast, Consents are less frequent, with 443 submissions (24.4%), likely due to their more specific requirements and potential complexity.
 
@@
 
-@tbl-application-type emphasize the dominance of Minor Variances in the dataset, providing context for their role in urban development.
+@tbl-application-type highlights the prevalence of Minor Variances in the dataset, providing context for their role in urban development.
 
 The relationship between application_type and the outcome variable decision will be explored further to determine if approval rates differ between Minor Variances and Consents.
-Preliminary trends suggest that Minor Variances may have higher approval rates due to their less complex nature.
+Preliminary trends indicate that Minor Variances may have higher approval rates, potentially due to their less complex nature.
 Additional analyses may examine whether these patterns vary by planning district or year.

### 2024-12-04 paper/paper.qmd
Paper Revision

@@
 The year variable represents the year of submission for each application in the dataset.
-It provides awareness into temporal trends in applications submitted to the Committee of Adjustment.
+It reflects temporal trends in applications submitted to the Committee of Adjustment.
 This variable is instrumental in identifying changes in the volume of applications over time, which may reflect policy shifts, urban development trends, or external factors like economic conditions.
@@
 
-@fig-year clarifies the distribution of applications from 2016 to 2024, with notable fluctuations in annual counts.
-Fewer applications were submitted during 2016-2019, with counts declining steadily from 198 applications in 2016 to a low of 93 in 2019.
+@fig-year illustrates the distribution of applications from 2016 to 2024, with notable fluctuations in annual counts.
+Fewer applications were submitted from 2016 to 2019, with a steady decline from 198 applications in 2016 to a low of 93 in 2019.
 This period may reflect a stabilization phase or lower activity in urban development processes.
-Afterwards, application counts increased significantly, peaking in 2023 with 378 applications, followed by a slight decline to 249 in 2024.
+Application counts then increased significantly, peaking at 378 applications in 2023, before declining slightly to 249 in 2024.
 The rise in 2023 may be due to heightened development activity or easing of restrictions from prior external events, such as the COVID-19 pandemic.
@@
 
-@tbl-year states how each year contributes to the total volume of applications.
+@tbl-year summarizes each year's contribution to the total volume of applications.
 
 The year variable is likely related to the decision outcome and application type.
-As approval rates may have varied over time due to changing city planning policies or committee priorities.
+Approval rates may have varied over time, potentially due to shifts in city planning policies or committee priorities.
 Besides, temporal trends may highlight shifts in the proportion of Minor Variances (MV) versus Consents (CO), reflecting evolving development demands.
@@
 The city is divided into four planning districts, each with its own development characteristics and zoning regulations.
-These districts are pivotal for understanding spatial trends in application submissions and outcomes, as local policies and priorities can influence decision-making processes.
+This variable helps identify spatial trends in application submissions and outcomes, as local policies and priorities can influence decision-making processes.
 
@@
 @fig-district displays the distribution of applications across four planning districts.
-We can observe that North York has the highest number of applications, accounting for a significant proportion of the dataset.
+North York has the highest number of applications, representing a significant proportion of the dataset.
 This could reflect the area's high density of development projects or active property owners seeking zoning adjustments.
@@
 
-@tbl-district provide a view of how applications are distributed spatially across the city.
-
-The planning_district variable may also interacts with application type and decision.
+@tbl-district summarizes how applications are distributed spatially across the city.
+
+The planning district variable may also interact with application type and decision.
 Approval rates may vary by district, reflecting differences in local policies or development dynamics.
@@
 
-In this paper, a Bayesian logistic model will be utilized to understand how factors such as application type, submission year, and planning district influence the decisions of the application committee.
+In this paper, a Bayesian logistic model is employed to understand how factors such as application type, submission year, and planning district influence the decisions of the application committee.
 This model estimates the probability of an application being approved (P(Approval)) based on the given predictors.
@@
 The goal of this modeling strategy is twofold: (1) to estimate how each predictor affects the likelihood of approval, and (2) to quantify the uncertainty associated with these estimates.
-This approach is well-suited for incorporating prior knowledge while addressing the hierarchical nature of the predictors, such as geographic regions.
+This approach accommodates prior knowledge while addressing the hierarchical nature of the predictors, such as geographic regions.
 
@@
 
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+Additional details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-In the above Bayesian logistic regression model, the coefficients $\beta = {\beta_0, \beta_1, \beta_2, \beta_3}$ are treated as random variables, each following a specified prior distribution.
-These priors represent our initial beliefs about the plausible values of the coefficients before observing the data.
-In this model, each $\beta_k$ (where k=0,1,2,3) is assigned a normal distribution, with a zero mean and 2.5 standard deviation.
-The choice of these parameters reflects a weakly informative prior, centered around zero to express no strong prior belief about the direction or magnitude of the effects, while allowing for reasonable variability in the coefficient estimates.
-
-The normal distribution is particularly suitable for these priors because it reflects a belief that most effects are likely to be small or moderate, centered around zero, while allowing for deviations in either direction.
-The standard deviation of 2.5 is a practical choice for weakly informative priors, offering a balance between constraining the parameters and allowing the data to influence the estimates.
+In the above Bayesian logistic regression model, the coefficients $\beta = {\beta_0, \beta_1, \beta_2, \beta_3}$ are treated as random variables, each assigned a Normal prior distribution, with a zero mean and 2.5 standard deviation.
+These priors reflect weakly informative beliefs, expressing no strong assumption about the direction or magnitude of effects while allowing for reasonable variability in the estimates.
+
+The Normal distribution is appropriate for these priors as it assumes most effects are small or moderate, centered around zero, while permitting deviations in either direction.
+A standard deviation of 2.5 offers a balance between constraining the parameters and allowing the data to influence the estimates.
 This ensures that the model avoids overfitting, particularly when data is sparse or multicollinearity exists among predictors.
 
-After observing the data, the posterior distributions of $\beta$ are obtained by combining these priors with the likelihood of the observed outcomes.
-The resulting posterior distributions provide updated beliefs about the coefficients, incorporating both prior knowledge and evidence from the data.
-This approach allows for robust inference, offering a clear quantification of parameter uncertainty and supporting more detailed interpretations of the predictors' effects.
+After observing the data, the posterior distributions of $\beta$ are derived by combining the priors with the likelihood of the observed outcomes.
+These posterior distributions represent updated beliefs about the coefficients, integrating both prior knowledge with evidence from the data.
+This approach provides robust inference, quantifies parameter uncertainty, and supports detailed interpretations of the predictors' effects.
 
@@
 
-The Bayesian logistic regression model is an appropriate choice for analyzing how application type, submission date, and planning district influence the likelihood of application approval.
-This choice is driven by the binary nature of the response variable, the model's ability to incorporate prior knowledge, and its probabilistic framework, which quantifies uncertainty more effectively than frequentist alternatives.
-Logistic regression naturally constrains predicted probabilities to the [0,1] interval, ensuring interpretability as likelihoods of approval, while the Bayesian approach enhances flexibility and provides richer revelations through posterior distributions.
+The Bayesian logistic regression model is suitable for analyzing how application type, submission date, and planning district influence the likelihood of application approval.
+This choice is driven by the binary nature of the response variable, the model's ability to incorporate prior knowledge, and its probabilistic framework, which effectively quantifies uncertainty than alternatives.
+Logistic regression naturally constrains predicted probabilities to the [0,1] interval, ensuring interpretability as likelihoods of approval, while the Bayesian approach adds flexibility and provides richer revelations through posterior distributions.
 
@@
 These priors help stabilize the model, particularly in cases where data is sparse or multicollinear, without overwhelming the contribution of the observed data.
-Additionally, Bayesian methods yield full posterior distributions rather than single-point estimates, allowing for a clearer understanding of parameter uncertainty and a better capacity to incorporate uncertainty into decision-making processes.
+Additionally, Bayesian methods yield full posterior distributions rather than single-point estimates, offering a clearer view of parameter uncertainty and a better capacity to incorporate uncertainty into decision-making processes.
 These advantages make the Bayesian logistic regression model preferable to a standard GLM, which provides point estimates and relies on asymptotic approximations for inference.
 
-The model also stands out against simpler approaches, such as simple linear regression (SLR).
+The model also addresses limitations of simpler methods, such as simple linear regression (SLR).
 While SLR might be used for binary outcomes in certain contexts, it is theoretically inappropriate because it does not constrain predictions to the [0,1] interval.
@@
 Moreover, SLR assumes a linear relationship between predictors and the response variable, which is unsuitable for binary outcomes.
-The logit transformation in logistic regression, by contrast, ensures a proper probabilistic framework while preserving interpretability in terms of log-odds.
+The logit transformation in logistic regression, by contrast, ensures a proper probabilistic framework and preserves interpretability in terms of log-odds.
 
@@
 
-`Mostly Harmless Econometrics` (@angrist2009mostly) addresses how to handle variability in time-series and cross-sectional data, focusing on ensuring robust and interpretable results.
-The use of submission year and planning district as key predictors aligns with Angrist and Pischke's guidance on analyzing temporal and geographic variation.
-Their discussion on the importance of controlling for time trends and location-specific effects validates the inclusion of these variables in the Bayesian logistic regression model.
-
-In summary, Bayesian logistic regression provides a robust, clarified, and flexible framework for analyzing the factors influencing application decisions.
+`Mostly Harmless Econometrics` (@angrist2009mostly) emphasizes addressing variability in time-series and cross-sectional data, focusing on ensuring robust and interpretable results.
+The inclusion of submission year and planning district as predictors follows Angrist and Pischke's guidance on controlling for time trends and location-specific effects, validating their relevance in this model.
+
+In summary, Bayesian logistic regression offers a robust and transparent framework for analyzing the factors influencing application decisions.
 Its advantages over GLM and SLR include the ability to incorporate prior information, quantify uncertainty, and handle the non-linear nature of the binary response variable.
@@
 
-This Bayesian model relies on several assumptions that ensure the validity of its predictions and the transparency of its coefficients.
-While these assumptions are generally less restrictive than those for simpler models like linear regression, they remain important to the robustness of the results.
-The key assumptions are listed below:
-
-1. Binary Response Variable
+This Bayesian model relies on several assumptions to ensure the validity of its predictions and the interpretability of its coefficients.
+While these assumptions are less restrictive than those for simpler models like linear regression, they remain important to the robustness of the results.
+The key assumptions are as follows:
+
+1. **Binary Response Variable**
 The model assumes that the response variable ($y$) is binary, taking values of either 1 (approval) or 0 (refused).
-This assumption aligns with the nature of the data, as application decisions are dichotomous outcomes.
-Any deviations from binary coding would invalidate the logistic framework and require alternative modeling approaches.
-
-2. Independence of Observations:
-The model assumes that all observations are independent of one another.
-This is reasonable for application data, as each decision is typically made independently by the committee.
-However, if clustering or dependence exists (e.g., decisions within the same planning district are correlated), the model may need to incorporate hierarchical or random effects to address these dependencies.
-
-3. No Perfect Multicollinearity
-Logistic regression assumes that the predictors are not perfectly correlated, as this would prevent the model from estimating unique coefficients.
+This assumption aligns with the nature of the data, as application decisions are dichotomous.
+Deviations from binary coding would invalidate the logistic framework and require alternative modeling approaches.
+
+2. **Independence of Observations**
+The model assumes that all observations are independent.
+This is reasonable for application data, as decisions are typically made independently by the committee.
+However, if clustering or dependence exists (e.g.
+correlated decisions within the same planning district), the model may require hierarchical or random effects to account for these dependencies.
+
+3. **No Perfect Multicollinearity**
+Logistic regression assumes that the predictors are not perfectly correlated, as this would prevent the model from unique coefficients.
 For example, if a predictor is a linear combination of other predictors, the model would fail to converge.
-In this analysis, categorical variables of application type and planning district are appropriately encoded to avoid such issues.
-
-4. Proper Model Specification
-The model assumes that all relevant predictors are included and correctly specified.
+In this analysis, categorical variables of application type and planning district are properly encoded to avoid such issues.
+
+4. **Proper Model Specification**
+The model assumes that relevant predictors are included and correctly specified.
 Omitting important variables or including irrelevant ones could lead to biased estimates or reduced transparency.
-For example, excluding interaction terms when they are theoretically justified might result in incomplete understanding of the predictors' effects.
+For example, excluding interaction terms when they are theoretically justified might lead to an incomplete understanding of the predictors' effects.
 

### 2024-12-04 paper/paper.qmd
Paper Revision

@@
 
-Violations of the above assumptions can impact this model’s validity while severl solution are listed below:
-
-1. Non-linearity:
+Violations of the above assumptions can impact this model’s validity while several solutions are listed below:
+
+1. **Non-linearity**
 Adding polynomial terms or splines could address non-linear relationships if diagnostic checks suggest deviations.
 
-2. Dependence:
+2. **Dependence**
 If observations are not independent (e.g. decisions clustered by region), hierarchical models or random effects may be introduced.
 
-3. Multicollinearity:
-Variance inflation factors or correlation matrices can help identify multicollinearity issues, which can be resolved by removing or combining correlated predictors.
-
-By adhering to these assumptions and addressing potential violations, the Bayesian logistic regression model provides a robust and effective framework for analyzing the factors influencing application outcomes.
+3. **Multicollinearity**
+Variance inflation factors or correlation matrices can identify multicollinearity issues, which can be resolved by removing or combining correlated predictors.
+
+By adhering to these assumptions and addressing potential violations, the Bayesian logistic regression model offers a reliable framework for analyzing the factors influencing application outcomes.
 
@@
 
-To ensure the robustness and reliability of the Bayesian logistic regression model, a series of validation and diagnostic steps were conducted.
+To ensure the robustness and reliability of the Bayesian logistic regression model, a series of validation and diagnostic steps was conducted.
 Detailed plots and results, including posterior predictive checks (PPC), posterior vs.
 prior comparisons, and convergence diagnostics, are provided in the appendix for further examination.
-Below, the focus is on key validation method of **out of sampling**:
+Below, the focus is on key validation method of **out-of-sample testing**:
 
@@
 
-@tbl-accuracy presents the train and test accuracy.
+@tbl-accuracy shows the train and test accuracy.
 The model's training accuracy of 66.75% and test accuracy of 66.575% are quite close, indicating that the model generalizes reasonably well between the training and test datasets.
-However, the relatively low accuracy on both datasets suggests the presence of underfitting, where the model fails to capture the complexity of the data and adequately explain the variance in the outcomes.
-This could result from the model being too simplistic, with insufficient predictors or inadequate representation of the data's structure.
-Additionally, potential issues like noisy or unbalanced data, insufficient feature engineering, or a lack of higher-order interactions and non-linearities in the model could contribute to this underfitting.
+However, the relatively low accuracy on both datasets suggests underfitting, where the model fails to capture the complexity of the data and adequately explain the variance in the outcomes.
+This could result from the model being too simplistic, with insufficient predictors or inadequate representation of the data structure.
+Additional factors, such as noisy or unbalanced data, limited feature engineering, or a lack of higher-order interactions and non-linearities in the model could also contribute to this underfitting.
 
@@
 The result highlights how the predictors influence the likelihood of application approval while controlling for other variables.
-The coefficients, presented as log-odds, provide observations into both the direction and magnitude of each predictor’s impact on the approval decision.
-
-The intercept estimate of -0.06 represents the baseline (an application of CO subimited in 2016 at Etobicoke York) log-odds of approval when all predictors are at their reference levels.
+The coefficients, presented as log-odds, indicate both the direction and magnitude of each predictor’s impact on approval decisions.
+
+The intercept estimate of -0.06 represents the baseline log-odds of approval for a consent (CO) application submitted in 2016 in Etobicoke York, where all predictors are at their reference levels.
 This corresponds to a baseline probability close to 0.5, indicating nearly equal odds of approval or refusal in the absence of other influencing factors.
@@
 
-The variable application_typeMV, with a coefficient of -0.20, suggests that applications for minor variances (MV) are less likely to be approved compared to consents (CO), which act as the reference category.
-In terms of odds, the transition from a consent to a minor variance reduces the odds of approval by approximately 18% ($e^{-0.20}\approx 0.82$).
-This finding aligns with expectations, as minor variances may face stricter scrutiny due to their potentially greater impact on local zoning adjustments.
-
-The year variables display significant temporal variations in approval likelihood, using 2016 as the reference year.
+The variable application_typeMV, with a coefficient of -0.20, suggests that applications for minor variances (MV) are less likely to be approved than consents (CO).
+In terms of odds, the transition from a consent to a minor variance reduces approval odds by approximately 18% ($e^{-0.20}\approx 0.82$).
+This aligns with expectations, as minor variances may face stricter scrutiny due to their potentially greater impact on local zoning adjustments.
+
+The year variables demonstrate significant temporal variations in approval likelihood, using 2016 as the reference year.
 These coefficients highlight how approval dynamics have shifted over time.
-For instance, in 2017, the coefficient of 0.74 indicates a 109% increase in the odds of approval compared to 2016 ($e^{0.74} \approx 2.09$), which suggests a substantial policy or procedural shift.
+In 2017, the coefficient of 0.74 indicates a 109% increase in the odds of approval compared to 2016 ($e^{0.74} \approx 2.09$), which suggests a substantial policy or procedural shift.
 In 2018, the strongest positive effect is observed, with a coefficient of 0.96, reflecting a 161% increase in odds relative to 2016 ($e^{0.96} \approx 2.61$).
-However, in 2019, the coefficient drops to 0.22, representing a smaller increase of 25% in odds ($e^{0.22} \approx 1.25$), suggesting that this period may have faced stricter reviews or procedural adjustments.
-By 2020, the coefficient rebounds to 0.56, translating to a 75% increase in odds ($e^{0.56} \approx 1.75$).
-
-For more recent years, 2021 sees a coefficient of 0.34, showing a 40% increase in odds ($e^{0.34} \approx 1.40$).
+However, in 2019, the coefficient drops to 0.22, representing a smaller increase of 25% ($e^{0.22} \approx 1.25$), suggesting that this period may have faced stricter reviews or procedural adjustments.
+By 2020, the coefficient rebounds to 0.56, translating to a 75% increase ($e^{0.56} \approx 1.75$).
+
+For more recent years, 2021 shows a coefficient of 0.34, showing a 40% increase in odds ($e^{0.34} \approx 1.40$).
 In 2022, the coefficient is 0.28, reflecting a 32% increase ($e^{0.28} \approx 1.32$).
@@
 Finally, 2024 showed the highest increase, with a coefficient of 1.24, translating to a 245% increase in approval odds compared to the baseline year of 2016 ($e^{1.24} \approx 3.46$).
-This steady upward trend reflects systemic improvements in the approval process, possibly due to increased efficiency, evolving zoning policies, or external influences favoring development.
+This upward trend reflects improvements in the approval process, possibly due to increased efficiency, evolving zoning policies, or external influences favoring development.
 
 The model captures geographic variability in approval rates across Toronto’s planning districts.
-North York, with a coefficient of 0.41, shows a 50% increase in odds compared to the baseline district ($e^{0.41} \approx 1.50$), highlighting its role as a development hub and its alignment with urban growth priorities.
-Scarborough, on the other hand, has the lowest approval likelihood, with a coefficient of 0.16, indicating a modest 17% increase in odds ($e^{0.16} \approx 1.17$), suggesting challenges related to stricter zoning rules or community resistance to changes.
+North York, with a coefficient of 0.41, shows a 50% increase in odds compared to Etobicoke York ($e^{0.41} \approx 1.50$), highlighting its role as a development hub with urban growth priorities.
+Scarborough has the lowest approval likelihood, with a coefficient of 0.16, indicating a modest 17% increase in odds ($e^{0.16} \approx 1.17$), suggesting challenges related to stricter zoning rules or community resistance to changes.
 Toronto East York has a coefficient of 0.33, showing a 39% increase in odds ($e^{0.33} \approx 1.39$), reflecting its mix of heritage preservation and urban development priorities.
-The results emphasize that approval rates vary significantly across districts due to localized zoning policies, development pressures, and socio-political factors influencing decision-making.
-
-In summary, the model presents key patterns in how application type and submission year influence the likelihood of approval.
-Minor variance applications face slightly lower odds of approval compared to consents, while temporal trends indicate an overall increase in approval rates from 2016 to 2020.
-These findings provide a foundation for understanding the dynamics of approval decisions and highlight areas for further investigation, such as geographic variability or interactions between predictors.
+The results emphasize localized influences on approval rates due to zoning policies, development pressures, and socio-political factors.
+
+In summary, the model displays how application type and submission year influence the approval likelihood.
+Minor variance applications face slightly lower odds of approval compared to consents, while temporal trends indicate an overall increase in approval rates from 2016 to 2024.
+Geographic variability further highlights differences across planning districts.
+These findings provide a basis for examining factors such as localized dynamics or interactions between predictors.
 
@@
 Consent applications exhibit a slightly higher median predicted probability of approval compared to minor variances.
-The interquartile range (IQR) for both application types is comparable, indicating a similar spread of probabilities.
-However, minor variance applications show a broader lower whisker, suggesting that certain cases have substantially lower probabilities of approval.
-This supports the interpretation that minor variances may face stricter scrutiny due to their greater potential impact on zoning.
+Both application types have similar interquartile ranges (IQR), but minor variances exhibit a broader lower whisker, reflecting stricter scrutiny for certain cases.
 
@@
 
-@fig-pp-district evaluates geographic variability in approval probabilities across the four planning districts: Etobicoke York, North York, Scarborough, and Toronto East York.
-North York consistently shows the highest median predicted probability of approval, reaffirming its status as a development hub with favorable zoning outcomes.
-In contrast, Scarborough exhibits the lowest median and a wider spread of probabilities, with outliers indicating challenging cases.
+@fig-pp-district evaluates geographic variability in approval probabilities across the four planning districts.
+North York consistently shows the highest median predicted probability of approval, reaffirming its development hub with favorable zoning outcomes.
+Scarborough has the lowest median and a wider spread of probabilities, with outliers indicating challenging cases.
 Toronto East York and Etobicoke York present intermediate medians, suggesting more balanced approval probabilities in these regions.

### 2024-12-04 paper/paper.qmd
Final Revision

@@
 This paper investigates the factors influencing application approvals submitted to the City of Toronto's Committee of Adjustment, using a Bayesian logistic regression model.
-By analyzing predictors such as application type, submission year, and planning district, the study provides findings into the dynamics of urban planning decisions.
+By analyzing predictors such as application type, submission year, and planning district, the study examines the dynamics of urban planning decisions.
 The model captures temporal trends, spatial variability, and procedural differences, offering a structured understanding of approval outcomes.
-Visualizations, including coefficient plots and effect plots, complement the analysis, helping to interpret the relationships between predictors and approval likelihood.
+Visualizations, including coefficient plots and predicted probability plots, complement the analysis, helping to interpret the relationships between predictors and approval likelihood.
 
@@
 One key finding is the temporal trend in approval rates.
-The analysis exhibits that approval odds have increased steadily over time, peaking in 2018.
+The analysis shows that approval odds have increased steadily over time, peaking in 2018.
 This trend may reflect evolving planning policies, greater efficiency in committee processes, or a shift in zoning priorities.
 The data suggests that the Committee of Adjustment has become more permissive in granting approvals, which could indicate a growing alignment between applications and city planning objectives.
-This result highlights how administrative processes adapt over time to urban development needs.
+This result demonstrates how administrative processes adapt over time to urban development needs.
 
@@
 
-While this paper provides findings into the factors influencing application approvals, there exists several limitations.
-These weaknesses cover data-related issues, methodological constraints, and broader contextual considerations.
+While this paper analyzes the factors influencing application approvals, several limitations remain.
+These weaknesses include data-related issues, methodological constraints, and broader contextual considerations.
 
@@
 The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain important variables.
-Key contextual factors, such as community objections, project complexity, or the presence of political or economic pressures, are not captured.
+Contextual factors, such as community objections, project complexity, or political and economic pressures, are not captured.
 For instance, the dataset does not provide information about why certain applications are refused or approved beyond the recorded decision, limiting the ability to analyze qualitative or situational influences.
@@
 This assumption may oversimplify real-world interactions between variables, such as how specific planning districts may interact with application types or temporal trends.
-The model's relatively low accuracy for both the training and test datasets suggests underfitting, indicating that it fails to capture the full complexity of the data and may not adequately explain the variance in outcomes.
-This limitation could stem from insufficient predictors, unaccounted non-linear relationships, or interactions among variables.
+The model's relatively low accuracy for both the training and test datasets suggests underfitting, failing to capture the data complexity and may not adequately explain the variance in outcomes.
+This could stem from insufficient predictors, unaccounted non-linear relationships, or interactions among variables.
 Furthermore, the model does not fully account for potential multicollinearity among predictors, which could distort the estimated coefficients.

### 2024-12-04 paper/paper.qmd
Final Revision

@@
 
-Our data [@committee_application] was extracted from opendatatoronto [@opendatatoronto].
+The data about Committee of Adjustment Applications [@committee_application] was extracted from Open Data Toronto [@opendatatoronto].
 This dataset focuses on applications submitted to the City of Toronto's Committee of Adjustment, a governing body responsible for reviewing minor variance and consent applications related to land use and development.
@@
   theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 0.5),
+    text = element_text(size = 6)
+  ) +
+  guides(color = guide_legend(ncol = 6)) +
   theme(legend.position = "bottom") +

### 2024-12-04 paper/paper.qmd
Paper Revision

@@
 
-`Sampling: Design and Analysis` (@lohr2021sampling) explores how stratified sampling reduces variability by dividing populations into homogeneous groups, enabling more precise estimates and better resource allocation.
+`Sampling: Design and Analysis` [@lohr2021sampling] explores how stratified sampling reduces variability by dividing populations into homogeneous groups, enabling more precise estimates and better resource allocation.
 By applying these principles, the paper ensures that the analysis accounts for heterogeneity in geographic and temporal patterns of planning applications, enabling robust conclusions about decision-making trends.
@@
 
-Besides, `Sampling` (@thompson2012sampling) provides a detailed guide to modern sampling techniques, emphasizing the importance of tailoring approaches to study objectives.
+Besides, `Sampling` [@thompson2012sampling] provides a detailed guide to modern sampling techniques, emphasizing the importance of tailoring approaches to study objectives.
 His discussion on stratified sampling highlights its benefits in improving precision and reducing bias by dividing populations into homogeneous subgroups.

### 2024-12-04 paper/paper.qmd
Final Revision

@@
 This complexity increases the time and resources required for sample design and execution.
-For instance, ensuring proportional representation across strata like **Scarborough developers** or **North York consultants** may necessitate oversampling in smaller groups, leading to additional effort in recruitment and logistical coordination.
+For instance, ensuring proportional representation across strata like Scarborough developers or North York consultants may necessitate oversampling in smaller groups, leading to additional effort in recruitment and logistical coordination.
 
 Another trade-off lies in the potential for overstratification, which can result in small sample sizes within certain strata.
-If some strata, such as **Scarborough Consent applications**, have very few members, it may be difficult to draw statistically meaningful conclusions.
+If some strata, such as Scarborough Consent applications, have very few members, it may be difficult to draw statistically meaningful conclusions.
 This could necessitate adjustments, such as combining smaller strata or applying post-stratification weighting, which could introduce bias or reduce interpretability.
@@
 Matching survey-reported outcomes with recorded decisions.
-Verifying consistency between applicant-reported application details and dataset attributes (e.g., application type, planning district).
+Verifying consistency between applicant-reported application details and dataset attributes.
 To ensure representativeness, response rates would be monitored, and additional recruitment efforts would target underrepresented groups or districts.
@@
 
-- >10 years
+- more than 10 years
 
@@
 
-Were there any external factors (e.g., economic conditions, policy changes) that influenced the decision?
+Were there any external factors (e.g. economic conditions, policy changes) that influenced the decision?
 
@@
 
-d[answer box]
+[answer box]
 
